---
title: "A Diamond in the Rough (Part b)"
output: pdf_document
---

Unit 1 - Lab 7b

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(mosaic)
require(plyr)
load("~/Dropbox/IDS Curriculum Design/Labs/Unit 1 Labs/Data/atus_dirty.RData")
```

Last time ...
=============

- We loaded our **American Time Use Survey** data and found that it had lots of problems.
    - The variable **names** weren't very descriptive.
    - Our _numerical variables_ were miss-specified as **strings** or **characters**
- **Explain the difference between the _string_ `r "118"` and the _number_ `r 118`.**


How did we fix these problems?
==============================

- We Loaded our data:
```{r, eval=FALSE}
data(atus_dirty)
```
- **Run this line of code**


- We renamed our variables
```{r, results='hide'}
names(atus_dirty) <- c("caseid",
                       "age",
                       "gender",
                       "fulltime_emp",
                       "phys_challenge",
                       "sleep",
                       "homework",
                       "socializing")
```
- **Run this line of code**



- We changed our **string** numbers back into **numerical** numbers:
```{r, results='hide'}
atus_dirty <- transform(atus_dirty,
                        age = as.numeric(age),
                        sleep = as.numeric(sleep),
                        homework = as.numeric(homework),
                        socializing = as.numeric(socializing))
```
- **Run this line of code**


So what's next?
===============

- Let's take a look at our data to find our next problem
```{r, eval=FALSE}
View(atus_dirty)
```

- **What do you notice about the `gender` and `phys_challenge` variables?**
- Recall that the variables tell us:
    - `gender`: The gender of the respondent.
    - `phys_challenge`: Whether the person has a physical difficulty.


Deciphering Categorical Variables
=================================

- Clearly, **gender** is a categorical variable but it's categories are represented by numbers.
    - This isn't necessarily a huge problem, but our data would be much clearer if we could replace the numbers `"01"` and `"02"` with `"Male"` and `"Female"`.
- The sames is true of the **phys_challenge** variables.

Factors and Levels
==================

- R has a special name for _categorical_ variables, called **Factors**.
- R also has a special name for the different _categories_ of a _categorical_ variable.
    - The individual categories are called **`"levels"`**.
- To see the levels of **gender** type:
```{r, eval=FALSE}
with(atus_dirty, levels(gender))
```


What's with with()?
===================

```{r, eval=FALSE}
with(atus_dirty, levels(gender))
```

- This line of code says:
    - "With our atus_dirty data..."
    - "... print out the levels of..."
    - "... the variable `gender`."

What's with() phys_challenge?
=============================

- **Using the method from the last slide, write down the `levels` of the `phys_challenge` variable**.


A level by any other name...
===========================

- If we know that '`01`' means '`Male`' and '`02`' means '`female`' then we can use the following code to revalue the **`levels`** of _gender_.
- Type the following command into your console:
```{r, results='hide'}
atus_dirty <- transform(atus_dirty, gender = revalue(gender, c("01" = "Male", "02" = "Female")))
```

- This code is definitely a bit of a mouthful. Let's break it down.

Allow me to explain
===================

```{r, results='hide'}
atus_dirty <- transform(atus_dirty, gender = revalue(gender, c("01" = "Male", "02" = "Female")))
```

- This code is saying:
    - "Replace my current version of `atus_dirty`..."
    - "... with a transformed one where ..."
    - "... the `gender` variables levels ..."
    - "... have been revalued..."
    - "... where '01' will now be 'Male'..."
    - "... and '02' will now be 'Female'."

Factors and Levels
==================

- `View` your data again and look at the values for **gender**
- **Rename the values of the variable `phys_challenge` where**
    - **`01:** No difficulty 
    - **`02:** Has difficulty

Ta-da!
======

- It took some work, but you should have a data set you can be proud of.
- Let's rename our data now that it's clean:
```{r,eval=FALSE}
atus_clean <- atus_dirty
```
- And let's also take a moment to admire it:
```{r, eval=FALSE}
View(atus_clean)
```



