---
title: "Oh the summaries..."
output: pdf_document
---



Unit 2 - Lab 2

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.
  
```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(MobilizeSimple)
```


Just the beginning
=====================

- Means, medians, MAD are just a few examples of **numerical summaries**
- **Numerical summaries** are numbers that describe characteristics of the data.
    - Means and medians described the _center_ of the data.
    - MAD describes the _spread_ of the data.
- **What other numbers might describe your data?**

To start
========

- Load your personality color data again and name it: ` colors`.
- In the lines of code that appear on the following slides:
    - Replace any of the `~x`  that appear in the code with the name of the variable for your _predominant_ color's score.

Extreme values
==================

- Besides looking at _typical_ values, sometimes we want to see _extreme_ values. Like the smallest and largest values.
- To find these values, we can calculate the `min` and `max`.
```{r,eval=FALSE}
min(~x, data=colors)
```
```{r,eval=FALSE}
max(~x, data=colors)
```

Range
=====

- Using the `min` and `max` values, we can calculate the `range`.
- The range is another, but often less informative, measure of _spread_.
    - Calculate it by taking: ` max - min`.

Calculating the range
=====================

- Try the following    
```{r,eval=FALSE}
max(~x, data = colors) - min(~x, data = colors)
```
```{r,eval=FALSE}
range(~x, data = colors)
```

- Notice how we can treat `max(~x, data = colors)` and `min(~x, data = colors)` just like we would a number.
- **Why did we say that the `range` is less informative than the `MAD`?** 
    - **Can you think of examples where the `MAD` will give you a better idea of the _variability_ than the `range`?**


Quartiles (Q1 & Q3)
=========================

- We often use the `median` to describe the _center_ of our data because half of the data is smaller than the median and the other half is larger.
- If instead we found a value that was larger than just 25% of our data, we would have computed the _1st quartile_.
- If we found the value that was larger than 75% of our data, we call that the _3rd quartile_.
```{r,eval=FALSE}

```
```{r,eval=FALSE}

```
- **Why do you think we use the names _`1st_ and _3rd quartiles_?**

The Inter-Quartile-Range (IQR)
========================

- Just like we used the `min` and `max` to comupute the `range`, we can also use the _1st_ and _3rd_ quartiles to compute the `IQR`.
- The `IQR` is another way to describe _spread_.
    - It describes how _wide_ or _narrow_ the middle 50% of our data are.
    - If the `IQR` is a small number, then the middle 50% of our data is close to the `median`.
    - Otherwise, the middle 50% of our data is further away from our median.
    
Finding the IQR
===============

- Make a histogram of your predominant color's scores.
- Visually: 
    - Cut the distribution into quarters so the _number_ of _data points_ is equal for each piece. (Each piece should contain 25% of the data.)
    - **Write down the numbers that split the data up into these 4 pieces.**
    - **How long is the interval of the middle two pieces?**
    - This length is the _IQR_.

Calculating the IQR
===================

- Calculate the IQR by using either of the following
```{r,eval=FALSE}
IQR(~x, data = colors)
```
```{r,eval=FALSE}
iqr(~x, data = colors)
```

- **How close was your visual estimate to the actual IQR?**

Other quantiles
===============

- The median, 1st and 3rd _quartiles_ can also be called the the 50th, 25th and 75th _quantiles_. 
    - They're called _quantiles_ becuase they're the _quantity_ of data that is smaller than that value.
    - The 25th quantile is the value that is larger than 25% of the data.
- We can compute quantiles too!
```{r,eval=FALSE}
qdata(~x, data = colors, p = 0.35)
```

- Where `p` stands for the _percentage_ of data you'd like our value to be larger than.

Boxplots
========

- By using the medians, quartiles and min/max, we can construct a new single variable plot called the **box and whisker** plot, often shortened to just a **boxplot**.
- Try making one of your predominant color.

```{r, eval=FALSE}
bwplot(~x, data=colors)
```

- **Sketch your boxplot in your journal. Label the `min` `max`, `Q1`, `Q3` and `median`.**
- **How would you interpret your boxplot? Where is the bulk of your data? Where is it centered? Can you say anything about its shape?**

Our favorite summaries
=====================

- Numerical summaries are brief ways to describe our data, using numbers.
- Computing lots of different summaries though can be tedious.
- Use the following command to compute some of our _favorite_ summaries
```{r, eval=FALSE}
favstats(~x, data=colors)
```
- **Which summaries are displayed?**
- **What do you think `n` stands for?**





