---
title: "Are you sure about that?"
author: "Lab 3D"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Confidence and intervals

- Throughout the year, we've seen that:
    - Means are used for describing the typical value in a sample or population, but we usually don't know what they are, because we can't see the entire population.
    - Means of samples can be used to _estimate_ means of populations.
    - By including a margin of error with our estimate, we create an interval that increases our confidence that we've located the correct value of the population mean.
- Today, we'll learn how we can calculate _margins of error_ by using a method called the _bootstrap_.
    - Which comes from the phrase, _Picking yourself up by your own bootstraps_.


## In this lab

- <span class="blue">Load the built-in `atus` (_American Time Use Survey_) dataset, which is a survey of how a sample of Americans spent their day.</span>
    - **The United States has an estimated population of 336,302,171 (as of April 15, 2024 9:10 a.m. PDT). How many people were surveyed for this particular dataset?**
    
    - Note: If you want to know the current US population or world population in real time, click on this link: [https://www.census.gov/popclock/](https://www.census.gov/popclock/)

- The statistical investigative question we wish to answer is:

_What is the mean age of people older than 15 living in the United States?_

- **Why is it important that the ATUS is a random sample?**
- **Use our `atus` data to calculate an estimate for the average age of people older than 15 living in the U.S.**


## One bootstrap

- A _bootstrapped_ sample is when we take a random `sample()` of our original data (`atus`) _WITH_ replacement.
    - The `size` of the sample should be the same size as the original data.
- We can create a single _bootstrapped_ sample for the `mean` in three steps:
    1. Sample the number of the rows to use in our _bootstrap_.
    2. `slice` those rows from our original data into our _bootstrap_ data.
    3. Calculate the mean our our _bootstrapped_ data.
    

## Our first bootstrap

- <span class="blue">Fill in the blanks to `sample` the row numbers we'll use in our _bootstrapped_ sample.</span>
    - Be sure to re-read what a _bootstrapped_ sample is from the previous slide to help you fill in the blanks.
    - Use `set.seed(123)` before taking the sample.

```{r, eval = FALSE, echo = TRUE}
bs_rows <- ____(1:____, size = ____, replace = ____)
```

- <span class="blue">Use the `slice` function to create a new dataset that includes each row from our `sample`.</span>

```{r, eval = FALSE, echo = TRUE}
bs_atus <- slice(atus, bs_rows)
```


## Take a look

- Look at the values of `bs_rows` and `bs_atus`.
- **Write a paragraph that explains to someone that's not familiar with `R` how you created `bs_rows` and `bs_atus`. Be sure to include an explanation of what the _values_ of `bs_rows` mean and how those values are used to create `bs_atus`. Also, be sure to explain what each argument of each function does.**


## One strap, two strap

- <span class="blue">Calculate the `mean` of the `age` variable in your `bootstrapped` data, then use a different value of `set.seed()` to create your own, personal _bootstrapped_ sample. Then calculate its `mean`.</span>
- **Compare this second _bootstrapped_ sample with three other classmates and write a sentence about how similar or different the _bootstrapped_ sample means were.**


## Many bootstraps

- To use _bootstrapped_ samples to create _confidence intervals_, we need to create many _bootstrapped_ samples.
    - Normally, the more _bootstrapped_ samples we use, the better the _confidence interval_.
    - In this lab, we'll `do()` 500 _bootstrapped_ samples.
- To make `do()`-ing 500 _bootstraps_ easier, we'll code our 3-step bootstrap method into a function.
    - <span class="blue">Open a new R Script (File -> New File -> R Script) to write your function into.</span>
    

## Bootstrap function

- <span class="blue">Fill in the blank space below with the 3 steps needed to create a _bootstrapped_ sample `mean` for our `atus` data.</span>
    - Each step should be written on its own line between the curly braces.

```{r, eval = FALSE, echo = TRUE}
bs_func <- function() {
    
    
    
}
```

- <span class="blue">Highlight and _Run_ the code you write. </span>


## Visualizing our bootstraps

- <span class="blue">Once your function is created, fill in the blanks to create 500 _bootstrapped_ sample means:</span>

```{r, eval = FALSE, echo = TRUE}
bs_means <- do(____) * bs_func()
```

- **Create a `histogram` for your bootstrapped samples and describe the _center_, _shape_ and _spread_ of its distribution.**
    - These bootstrapped estimates no longer estimate the average age of people in the U.S.
    - Instead, they estimate how much the estimate of the average age of people in the U.S. varies.
- In the next slide, we'll look at how we can use these bootstrapped means to create _90% confidence intervals_.


## Bootstrapped confidence intervals
    
- To create a 90% confidence interval, we need to decide between which two _ages_ the middle 90% of our bootstrapped estimates are contained.
- **Using your `histogram`, fill in the statement below:**

The lowest 5% of our estimates are below _______ years and the highest 5% of our estimates are above_______ years.

- <span class="blue">Use the `quantile()` function to check your estimates.</span>
- **Based on your bootstrapped estimates, between which two ages are we 90% confident the actual `mean` age of people living in the U.S. is contained?**


## On your own

- <span class="blue">Using your _bootstrapped_ sample means, create a 95% confidence interval for the `mean` age of people living in the U.S.</span>
- **Why is the 95% confidence interval wider than the 90% interval?**
- **Write down how you would explain what a 95% confidence interval means to someone not taking _Introduction to Data Science_.**