---
title: "Getting it together"
output: pdf_document
---


Unit 2 - Lab 8  

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.
 
```{r, echo=FALSE, message=FALSE}
require(MobilizeSimple)
```


Putting data together
=====================

- In this lab, we’ll continue our investigation into whether the personality color traits are meaningful. 
    - We'll investigate whether your personality color says anything about how stressed or chill you are during the day.  
    - To do this, we’ll need to merge two datasets together.
- Load your personality color data and name it `colors`.
- Then _Download, upload, load_ your stress/chill data set and call it `stress`.


Looking at stress/chill
=======================

- Go ahead and `View` the stress/chill data.
- Based on the variables in the stress/chill data:
    - **Which of the provided variables do you think might tend to raise or lower people's stress/chill level?**
    - **Do you think people's personalities might effect their general level of stress?** 
    - **Using just the stress/chill data, could you answer the question above?**


    
Deciding how to merge
=====================

- Before we merge data, we need to decide _how_ we plan on merging it.
    - We can _stack_ our data sets, that is take one data set's rows and add them to the bottom of the other.
    - We can also _join_ data sets horizontally. Where we take take one data set's columns and take them on at the end of the other data set's columns.
- We have data about people's stress levels in our `stress` data and data about people's personalities in our `colors` data.
    - **Write down the steps you would take if you had to merge these data sets together manually**

Finding variables in common:
============================

- Look at the `names` of the variables in each data set.
- To merge different data sets together, we need to find variables they have in common.
    - **Which variables do the data sets have in common?**
    - **Which variable would make sense to merge the data sets together with? Why not one of the others?**
    

Putting them together
=====================

- Before merging the variables together:
    - **Write down the dimensions of each data set**
    - **Wride down the number of rows and columns you think the _merged_ data should have.**
- To merge the data sets together, run the following command:
```{r,eval=FALSE}
merge(stress, colors, by = "user.id")
```
- Save this `merged` data set as `stress_colors`.
    
Looking at our data
====================
- `View` your merged data. Does it look _right_?
- **What are the dimensions of this new `stress_colors` data set?**
    - **Why did it make sense to combine the data by joining the columns?** 
    - **Why didn't we stack the rows of data instead?**
- **What happens if you swap the order of the data sets in the code above? Do the dimensions change?**
    

Saving our file to use later
============================

- Now that we've created the data set we want, it makes sense for us to save it for later use.
- Run the following code to save your  `stress_colors` data:
```{r,eval=FALSE}
save(stress_colors, file = "~/stress_colors.rda")
```

- Look in your home folder in the _Files_ tab to make sure your data was saved.


A preview...
============

- In the next lab, we'll analyze our newly merged data. But for now, try creating a plot to answer the following question:
    - **Did some (predominant) personality colors spent more time outside than others? If so, which? Write down why you think that might be the case.**



