---
title: 'What Stresses Us?'
output: pdf_document
---

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.
```{r, echo=FALSE, message=FALSE}
require(MobilizeSimple)
```

In the previous lab...
======================

- We made a data set that combined our _stress/chill_ data with our _personality color_ data.
    - Load your data by clicking on it in the _Files_ pane.
- In case you missed the last lab (Or didn't save your merged data): 
    - Load your _personality color_ data and name it `colors`.
    - Load your _stress/chill_ data and name it `stress`.
    - Then run the following to merge them together:
```{r, eval=FALSE}
stress_colors <- merge(stress, colors, 
                       by = "user.id")
```


For this lab ...
=================

- We'll use the techniques we learned in the previous labs to explore and analyze our `stress_colors` data.


Stress/Chill
================

- Make a plot that shows the distribution of the `stresschill` values.
- Using only your plot:
    - **Describe the _shape_ of the distribution.**
    - **Typically, what `stresschill` level did your class support.**
    - **Estimate the _variability_ of `streechill` values.**
- **Write a sentence explaining how and why you chose your particular values to describe the _variablity_ and _center_ of the data.**


Stress/Chill & Sports
=====================

- Create two boxplots of your `stresschill` values based on whether the person played sports or not.
- Based on your plot:
    - **Does it appear that one group has higher levels of stress than the other? Justify your answer.**
    - **Compute (Don't estimate the answers using the plot) the _min_, _max_, _Q1_, _median_, _mean_ and _Q3_ for each group.** 
    - (HINT: You can compute these numbers with a single line of code. Check lab 2.2 if you forgot how.)

Colors & Sports
===============

- The color test predicts that _ORANGE_ people like physical activity, and so we should see more sports players in the orange than the other colors. 
- **Does the data support this claim?**
    - **Write out the code you used to determine this answer.**
- **Could it be that the proportion of sports players who are _ORANGE_ is just by chance? Answer this question by comparing the actual proportion of sports players who are _ORANGE_ to 300 randomized trials, answer the following question:**
    - **Justify your answer by comparing the _actual_ data to the 300 randomized trials.**