---
title: "Zooming Through Data"
author: "Lab 1D"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Data with Clarity

- Previously, we've looked at graphs of entire variables (by looking at all of their values).
    - Doing this is helpful to get a _big picture_ idea of our data.
- In this lab, we'll learn how to _zoom in_ on our data by learning how to subset.
    - We'll also learn a few ways to manipulate the plots we've been making to make them easier to use for analyses.
- Import the data from your class' _Food Habits_ campaign and name it `food`.

## Another plotting function

- A `dotPlot` is another plot that can be used to analyze a numerical variable.
  - Dotplots are better suited for smaller data sets. If data sets are too large, the dots become too small to see.
  - Similarly, distributions with a large spread might impact the readability of the plot.
- <span class="blue">Use the `dotPlot()` function to create a `dotPlot` of the amount of `sugar` in our `food` data.</span>
    - The code to create a `dotPlot` is exactly like you'd use to make a `histogram`.
    - Make sure to use a capital _P_ in `dotPlot`.
    

## More options

- While a `dotPlot` should conserve the exact value of each data point, sometimes it behaves like a `histogram` in that it lumps values together.

- <span class="blue">Create a more accurate `dotPlot` by including the `nint` option.</span>
  - Set `nint` equal to the maximum value for sugar minus the minimum value for sugar plus one.
    - <span class="blue">On your `food` data spreadsheet, click on the sugar header to sort in ascending order (to obtain minimum). </span>
    - <span class="blue">Click on the sugar header again to sort in descending order (to obtain maximum). </span>
  - <span class="blue">Use your history pane to see how we included the option `nint` with the `histogram` function.</span>
  
- Pro tip: If the `dotPlot` comes out looking wonky, try changing the value of the _character expansion_ option, `cex`. 
    - The default value is `1`. Try a few values between `0` and `1` and a few more values larger than `1`.

## Splitting data sets

- In lab 1B, we learned that we can _facet_ (or split) our data based on a categorical variable.

- <span class="blue">Split the `dotPlot` displaying the distribution of grams of sugar in two, by faceting on our observations' `salty_sweet` variable.</span>
    - **Describe how `R` decides which observations go into the left or right plot.**
    - **What does each _dot_ in the plot represent?**


## Altering the layout

- It would be much easier to compare the sugar levels of salty and sweet snacks if the dotPlots were stacked on top of one another.
- We can change the _layout_ of our separated plots by including the `layout` option in our `dotPlot` function. 
    - <span class="blue">Add the following option to the code you used to create the `dotPlot` split by `salty_sweet`.</span>
```{r, eval = FALSE, echo = TRUE }
layout = c(1,2)
```
- Hint: Use a similar syntax used with the `nint` option to add the `layout` option to the `dotPlot` function.

## Subsetting

- Subsetting is a term we use to describe the process of looking at only the data that conforms to some set of rules:
    - Geologists may subset earthquake data by looking at only large earthquakes.
    - Stock market traders may subset their trading data by looking only at the previous day's trades.
- There's _many_ ways to subset data using RStudio, we'll focus on learning the most common methods.  


## The filter function

- Creating a plot for `Salty` and a plot for `Sweet` is useful for comparing `Salty` and `Sweet`. What if we want to examine one group by itself?
- The line of code below creates a subset of the `food` dataset containing only `Salty` snacks. We will break it down piece by piece in the next few slides.
- <span class="blue">Run the line of code below.</span>

```{r, eval = FALSE, echo = TRUE}
food_salty <- filter(food, salty_sweet == "Salty")
```
- <span class="blue">`View` `food_salty` and write down the number of observations in it.</span>


## So what's really going on?

- Coding in `R` is really just about supplying directions in a way that `R` understands. 
    - We'll start by focusing on everything to the right of the `<-` symbol
```{r, eval = FALSE, echo = TRUE}
food_salty <- filter(food, salty_sweet == "Salty")
```
- `filter()` tells R that we're going to look at only the values in our data that follow a _rule_.
- The first blank should be the data we're going to filter down into a smaller set (based on our rule).
- `salty_sweet == "Salty"` is the rule to follow.


## 3 parts of defining rules

- We can decompose our rule, `salty_sweet == "Salty"`, into 3 parts:
    1. `salty_sweet`, is the particular _variable_ we want to use to select our subset.
    2. `"Salty"` is the _value_ of the variable that we want to select. We only want to see data with the value `"Salty"` for the variable `salty_sweet`.
    3. `==`  describes how we want to relate our variable (`salty_sweet`) to our value (`"Salty"`). In this case, we want values of `salty_sweet` that are _exactly equal_ to `"Salty"`.
- Notice: _Values_ (that are also words) have quotation marks around them. _Variables_ do not.


## More on ==

- We can use the `head()` function to help us see what's happening when we write `salty_sweet == "Salty"`.
    - `head()` returns the values of the first 6 observations.
    - The `tail()` function returns the last 6 observations.
- Run the following code and answer the question below:
```{r, eval = FALSE, echo = TRUE}
head(~salty_sweet == "Salty", data = food)
```
- **What do the values `TRUE` and `FALSE` tell us about how our _rule_ applies to the first six snacks in our data? Which of the first six observations were `Salty`?**

## Saving values

- To use our subset data we need to save it first.
    - When we _save_ something in `R` what we are really doing is giving a value, or set of values, a specific name for us to use later.
- The arrow <- is called the "assignment" operator. It assigns names (on the left) to values (on the right).
    - We now focus on everything to the left of, and including, the "<-" symbol.
```{r, eval = FALSE, echo = TRUE}
food_salty <- filter(food, salty_sweet == "Salty")
```


## Saving our subset

```{r, eval = FALSE, echo = TRUE}
food_salty <- filter(food, salty_sweet == "Salty")
```

- This code then: 
    - takes our subset data, (everything to the right of `<-`) ...
    - and assigns the subset data, by using the arrow `<-` ...
    - the name `food_salty`.
- We can now use `food_salty` to do anything we could do with the regular `food` data ...
    - but only including those snacks who reported being `Salty`.
- As a result of assigning the subset data to `food_salty`, `food_salty` now appears in the _Environment_ pane. Whenever data is assigned to a variable name, that variable name will appear in the _Environment_ pane.
- <span class="blue">Use `food_salty` to make a `dotPlot` of the `sodium` in our `Salty` snacks.</span>

## Including more filters

- We often want to filter our data based on multiple rules.
    - For instance, we might want to filter our `food` data based on the food being salty AND having less than 200 calories.
- We can include multiple filters to our subsets by separating each rule with a comma like so:

```{r, eval = FALSE, echo = TRUE}
my_sub <- filter(food , salty_sweet == "Salty", calories < 200)
```

- `View` the  `my_sub` data we filtered in the above line of code and verify that it only includes salty snacks that have less than 200 calories.


## Put it all together

- **Use an appropriate `dotPlot` to answer each of the following questions:**
    - **About how much `sugar` does the typical sweet snack have?**
    - **How does the typical amount of `sugar` compare when `healthy_level < 3` and when `healthy_level > 3`?**
    
- Because you are now working with subsets of data, it is important to label our plots and make this distinction.

  - We can use the `main` option to add a title to our plots.
    - <span class="blue">Add the following option to the code you used to create the `dotPlot` of the `sugar` in `Sweet` snacks.</span>
```{r, eval = FALSE, echo = TRUE }
main = "Distribution of sugar in sweet snacks"
```
