---
title: "What's the FREQ?"
output: pdf_document
---

Unit 1 - Lab 8

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(mosaic)
load("~/Dropbox/IDS Curriculum Design/Labs/Unit 1 Labs/Data/atus_clean.RData")
```


USA! USA!
=========

- In the previous lab, we cleaned our American Time Use Survey (ATUS) data to make it usable.
- Now we can actually start to analyze it to learn how people spend their time in the US.
- The process of cleaning and then analyzing data is very common in Data Science.



Summarizing data
================

- Plots are extremely useful.
    - Help find interesting patterns.
    - Easy to make comparisons.
- Load your cleaned ATUS data set:
```{r,eval=FALSE}
data(atus_clean)
```
- **Create a plot that compares people's genders and the amount of time they spent on homework**
    - **Using this plot, do you think males or females spend more time doing homework?**
    - **Find someone in your class who came to the opposite conclusion as you did about which gender, typically, does more homework**.

- Instead of relying solely on plots, we want to **make comparisons** using numbers.


How do we summarize categorical variables?
=================================

- When we're dealing with categorical variables, we can't just calculate an **average** to describe a _typical_ value.
    - (Honestly, what's the average of categories _orange_, _apple_ and _banana_, for instance?)
- When trying to describe categorical variables with numberws, we calculate **frequency tables**

Frequency tables?
=================

- When it comes to categories, about all you can do is **count** or **tally** how often each category comes up in the data.
- Let's calculate how many _males_ and _females_ are in our data set.
- Type the following into the console:
```{r,eval=FALSE}
tally(~gender, data = atus_clean)
```
- **How many more _females_ than _males_ are there?**

2-way Frequency Tables
======================

- Counting the categories of a single variable is nice, but often times we want to make comparisons.
    - For example, what if we wanted to compare the number of people with phyiscal challanges and their genders?
- To makes these types of comparisons, we can make a **2-way frequency table**:
```{r,eval=FALSE}
tally(phys_challenge~gender, data = atus_clean)
```
- **Run this command and write down what you notice about numbers of _males_ and _females_ with physical challenges. Are we still counting?**

Interpreting 2-way frequency tables
===================================

- Recall that  there were `r tally(~gender, data = atus_clean)[2] - tally(~gender, data = atus_clean)[1]` more women than men in our data set.
    - Comparing **counts** then doesn't make sense.
    - If there are more women, then we might  expect women to have more physical challenges (compared to men).
- Instead of using **counts**, we use **percentages**
    - So for instance, roughly `r tally(phys_challenge~gender, data = atus_clean)[1,1] * 100`% of men do not have a physical difficulty.
    
Why do we use percentages?
==========================

- Percentages let us make comparisons between groups, even if one group is much larger than another.
- Run these two lines of code:

```{r,eval=FALSE}
tally(phys_challenge~gender, data = atus_clean)
```

```{r,eval=FALSE}
tally(phys_challenge~gender, data = atus_clean)
```

- **How many _females_ had a _physical challenge__ (using counts)?**
- **How many _males_ had a _physical challenge__ (using counts)?**

Changing our format
===================

- If we did want to make a table with _counts_ instead of _percentages_, we can change the `format` of our frequency table.
- Run the following line:
```{r, eval=FALSE}
tally(phys_challenge~gender, data = atus_clean, format = 'count')
```
- **Which gender had more people report NOT having a physical challenge, in terms of counts?**
- **Which gender had more people report NOT having a physical challenge, in terms of percents?**

Adding margins
==============

- Making frequency tables with counts can be misleading.
- To make them less so, we can add in the **margins** or totals for each gender.
- Run this line of code:
```{r, eval=FALSE}
tally(phys_challenge~gender, data = atus_clean, 
      format = 'count', margins = TRUE)
```
- **Explain what happened to the output by including `margins = TRUE` in the function.**


On your own
===========

- **Which gender has a higher rate of _part time employment_?**
- **Explore the amount of time each gender socializes**
    - **Create a subset of data of people who socialized at least 1 minute or more**
    - **Compare the average amount of socializing done by each gender**
    - **Does one gender's amount of socializing vary more than the other?**




