---
title: "Our time."
author: "Lab 1H"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(mobilizr)
```

## We've come a long way

- The labs until now have covered a huge range of topics:
    - We've learned how to make plots for different types of variables.
    - We know how to subset our data to get a more refined view of our data.
    - We've covered cleaning data and making two-way frequency tables.
- In this lab, we're going to combine all of these ideas and topics together to find out how we spend our time.


## First steps first.

- <span class="blue">_Export_, _Upload_, _Import_ the data from your class' _Time Use_ campaign.</span>

- The data, as-is, is very messy and hard to interpret/analyze.

    - <span class="blue">Fill in the blank with the name of your imported data to format it:</span>
```{r, eval = FALSE, echo = TRUE}
timeuse <- timeuse_format( _____ )
```
- This function formats/cleans the data so that each row represents a typical day for each student in the class
- <span class="blue">Hint: Search your _History_ tab for the code to save your formatted `timeuse` data as an `R` data file (.Rda)</span>


## timeuse_format specifics

- In case you're wondering, the `timeuse_format` function:
    - Takes each student's daily data and adds up all of the time spent doing each activity for each day.
    - The time spent on each activity for each day is then averaged together to create a _typical day_ in the life of each student.


## Exploring your data

- Start by getting familiar with your `timeuse` data: 
    - **How many observations and variables are there?**
    - **What are the names of the variables?**
    - **Which row represents YOUR _typical day_?**
  
  
## How do we spend our time?

- We would like to investigate the _research question_: "How did our class spend our time?"
    - To do this, we'll perform a statistical investigation.
- **State and answer two investigative questions based on our _research question_.**  
    - **Also, state one way in which your personal data is _typical_ and one way that it _differs_ from the rest of the class.** 
- **Justify your answers by using appropriate statistical graphics and summary tables.** 
    - **If you subset your data, explain why and how it benefited your analysis.**