---
title: "What's the score?"
author: "Lab 4B"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Previously

- In the previous lab, we learned we could make predictions about one variable by utilizing the information of another.
- In this lab, we will learn how to measure the accuracy of our predictions.
    - This in turn will let us evaluate how well a model performs at making predictions.
    - We'll also use this information later to compare different models to find which model makes the best predictions.


## Predictions using a line

- Load the `arm_span` data again.
    - Create an `xyplot` with `height` on the y-axis and `armspan` on the x-axis.
    - Type `add_line()` to run the `add_line` function; you'll be prompted to click twice in the plot window to create a line that you think fits the data well.
- Fill in the blanks below to create a function that will make predictions of people's `height`s based on their `armspan`:
```{r, eval = FALSE, echo = TRUE}
predict_height <- function(armspan) {
  ____ * armspan + ____
}
```


## Make your predictions

- Fill in the blanks to include your predictions in the `arm_span` data.
```{r, eval = FALSE, echo = TRUE}
____ <- mutate(____, predicted_height = ____(____))
```
- Now that we've made our predictions, we'll need to figure out a way to decide how accurate our predictions are.
    - We'll want to compare our _predicted heights_ to the _actual heights_.
    - At the end, we'll want to come up with a single number summary that describes our model's accuracy.


## Sums of differences

- A _residual_ is the difference between the actual and predicted value of a quantity of interest.
- Fill in the blanks below to add a column of residuals to `arm_span`.

```{r, eval = FALSE, echo = TRUE}
____ <- mutate(____, residual = ____ - ____)
```

- **What do the residuals measure?**

- One method we might consider to measure our model's accuracy is to sum the residuals.
- Fill in the blanks below to calculate our accuracy summary.

```{r, eval = FALSE, echo = TRUE}
summarize(____, sum(____))
```

- Hint: Like `mutate`, the first argument of `summarize` is a dataframe, and the second argument is the action to perform on a column of the dataframe. Whereas the output of `mutate` is a column, the output of `summarize` is (usually) a single number summary.

## Checking our work

- **Describe and interpret, in words, what the output of your accuracy summary means.**
    - **Compare your accuracy summary with a neighbor. Whose line was more accurate and why?**
- **Write down why adding positive and negative errors together is problematic for accessing prediction accuracy.**
    - **Why does calculating the squared values for the differences solve this problem?**
    
- The _mean squared error_ (MSE) is calculated by squaring all of the residuals, and then taking the mean of the squared residuals.
- Fill in the blanks below to calculate the MSE of your line.

```{r, eval = FALSE, echo = TRUE}
summarize(____, mean((____)^2))
```

## On your own

- Create a _regression line_ as you did in the previous lab, for `height` and `armspan`.
    - We also refer to _regression lines_ as _linear models_.
    - Assign this model the name `best_fit`. 
- Making predictions with models `R` is familiar with is simpler than with lines, or models, we come up with ourselves.
    - Fill in the blanks to make predictions using `best_fit`:
```{r, eval = FALSE, echo = TRUE}
____ <- mutate(____, predicted_height = predict(____))
```

- Hint: the `predict` function takes a linear model as input, and outputs the predictions of that model.
- Calculate the MSE for these new predicted values.


## The magic of lm()

- The `lm()` function creates the _line of best fit_ equation by finding the line that minimizes the _mean squared error_. Meaning, it's the _best fitting line possible_. 
    - Compare the MSE value you calculated using the line you fitted with `add_line()` to the the same value you calculated using the `lm` function.
    - Ask your neighbors if any of their lines beat the `lm` line in terms of the MSE. Were any of them successful?
- To see how the `lm` line fits your data, create a scatterplot and then run:
```{r, eval = FALSE, echo = TRUE}
add_line(intercept = ____, slope = ____)
```

