---
title: "Lab 2H - Eyeballing Normal"
output: word_document
---


Directions: Follow along with the slides and answer the questions in **bold** font in your journal.

\newline

 
```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## What's normal?


- The _normal distribution_ is a curve we often see in real data.
    - We see it in people's blood pressures and in measurement errors.
- When data appears to be _normally distributed_, we can use the _normal model_ to:
  - Simulate _normally distributed_ data.
  - Easily compute probabilities.
- In this lab, we'll look at some previous data sets to see if we can find data that are roughly normally distributed.


## The normal distribution


- The normal distribution is _symmetric about the mean_:
    - The `mean` is found in the very center of the distribution.
    - And the curve looks the same to the left of the mean as it does on the right.
- Use the following to draw a normal distribution:
```{r, eval = FALSE}
plotDist('norm', mean = 0, sd = 1)
```


## The mean and sd of it


- To draw a normal curve, we need to know exactly 2 things:
    - The `mean` and `sd`.
- The `sd`, or _standard deviation_, is a measure of spread that's similar to the `MAD`.
- **Which part of the normal curve changes when the value of the `mean` changes?**
- **Which part of the normal curve changes when the value of the `sd` changes?**
- _Hint_: Try changing the `mean` and `sd` values in the `plotDist` function.


## Finding normal distributions


- Load the `cdc` data and use the `histogram` function to answer the following:
- **Based on what you know about these variables, which of the variables do you think have distributions that will look like the normal distribution?**
    - **Make histograms of these variables.  Which ones look like the normal distribution?**
    - _Hint_: To help answer this question, try including the option `fit = "normal"` in the histogram function. You might also try faceting by `gender`.


## Using normal models


- Data scientists like using normal models because it often resembles real data.
    - _But not EVERYTHING is normally distributed._
- As a data scientist in training, you must decide when a normal model seems appropriate.
    - No model is ever perfect 100% of the time.
    - If you choose a model, you should be able to justify why you chose it.


## On your own


- **For each of the following, determine which, if any, appear to be normally distributed. Explain your reasoning:**
    - **The `weight` of people in our `cdc` data, faceted by `gender`.**
    - **The difference in `mean` weights between `Male`s and `Female`s for 500 random shuffles.**
    - **The difference in `median` weights between `Male`s and `Female`s for 500 random shuffles.**

