---
title: "The results are in!"
author: "Lab 3A"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Conducting experiments

- Previously in class, you conducted an experiment to gauge how a stimulus affected people's perception of time.
    - Some people were given a treatment, others were not.
- In this lab, we'll use the data cycle to analyze the _research question_:

_Does the stimulus your class chose change people's perception of time?_


## Coming up with questions

- **Write down two statistical investigative questions that will help you answer the _research question_ from the previous slide.**
- <span class="blue">Then, _export_, _upload_, _import_ your experiment data into RStudio.</span>
    - If you're having trouble coming up with good statistical investigative questions, try loading the data and looking at the variables.
    - Ask yourself, _How would I use these variables to answer the research question?_


## Analyzing our data

- <span class="blue">Create appropriate plots to answer your statistical investigative questions.</span>
- **Are there any outliers or unusual observations that require some cleaning before you can interpret your plots?**
- <span class="blue">Calculate appropriate numerical summaries to answer your statistical investigative questions.</span>
- **Write down a few sentences interpreting your plots and summaries.**


## Wrapping it up

- Is it possible your initial results occurred by chance alone?
    - <span class="blue">Use repeated shuffling to determine how likely the typical difference between the two groups occurred by chance alone.</span>
    - **Create a plot and use it to justify your answer.**
- What do you conclude about the _research question_?
    - **Write a report using the plots and analysis you conducted to answer the _research question_.**
    - Be sure to describe how you conducted your experiment. 