---
title: "All About Distributions"
author: "Lab 2A"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## In the beginning ...

- Most of the labs thus far have covered how to visualize, summarize, and manipulate data.
    - We used visualizations to explore how your class spends their time.
    - We also learned how to clean data to prepare it for analyzing.
- Starting with this lab, we'll learn to use R to answer statistical questions that can be answered by calculating the mean, median and MAD.


## How to talk about data

- When we make plots of our data, we usually want to know:
  - Where is the _bulk_ of the data?
  - Where is the data more _sparse_, or _thin_?
  - What values are _typical_?
  - How much does the data _vary_?
- To answer these questions, we want to look at the _distribution_ of our data.
    - We describe _distributions_ by talking about where the _center_ of the data are, how _spread_ out the data are, and what sort of _shape_ the data has.


## Let's begin!

- _Export_, _upload_ and _import_ your class' _Personality Color_ data.
    - Name your data `colors` when you load it.
- Before analyzing a new data set, it's often helpful to get familiar with it. So:
    - **Write down the `names` of the 4 variables that contain the point-totals, or _scores_, for each personality color.**
    - **Write down the `names` of the variables that tell us an observation's introvert/extrovert designation and whether they are involved in _sports_.**
    - **How many variables are in the data set?**
    - **How many observations are in the data set?**


## Estimating centers

- Create a `dotPlot` of the scores for your _predominant color_.
    - Pro-tip: If the `dotPlot` comes out looking wonky, include the `nint` and `cex` options.
- Based on your `dotPlot`:
    - **Which values came up the most frequently? About how many people in  your class had a score similar to yours?**
    - **What, would you say, was a _typical_ score for a person in your class for your predominant color? How does your own score for this color compare?**


## Means and medians

- _Means_ and _medians_ are usually good ways to describe the _typical_ value of our data.
- Fill in the blank to calculate the `mean` value of your predominant color score: 

```{r, eval = FALSE, echo = TRUE}
mean(~____, data = colors)
```

- **Use a similar line of code to calculate the `median` value of _your_ predominant color.**
    - **Are the `mean` and `median` roughly the same? If not, use the `dotPlot` you made in the last slide to describe why.**

## Estimating Spread

- Now that we know how to describe our data's _typical_ value we might also like to describe how closely the rest of the data are to this _typical_ value. 
    - We often refer to this as the **variability** of the data. 
    - Variability is seen in a `histogram` or `dotPlot` as the horizontal *spread*.
    
- Re-create a `dotPlot` of the scores for your _predominant color_ and then run the code below filling in the blank with the name of your predominant color:

```{r, eval = FALSE, echo = TRUE}
add_line(vline = mean(~____, data = colors))
```

- **Look at the spread of the scores from the mean score then complete the sentence below:**

_Data points in my plot will usually fall within_ ____ _units of the center._

## Mean Absolute Deviation

- The **mean absolute deviation** finds how far away, on average, the data are from the mean.
    - We often write _mean absolute deviation_ as _MAD_.
- Calculate the MAD of your _predominant color_ by filling in the blank:

```{r, eval = FALSE, echo = TRUE}
MAD(~_____, data = colors)
```

- **How close was your estimate of the spread for your predominant color (from the previous slide) to the actual value?**

## Comparing introverts/extroverts

- Do introverts and extroverts differ in their typical scores for your predominant color?
  - Answer this investigative question using a dotPlot and numerical summaries.
- Make a `dotPlot` of your _predominant color_ again; but this time, facet the plot by the introvert/extrovert variable. Include the `layout` option to stack the plots as well as the `nint` and `cex` options.
- **Describe the shape of the distribution of scores for the extroverts. Do the same for the introverts.**
- Using similar syntax to how you facet plots, _calculate_ either the `mean` or `median` to describe the _center_ of your predominant color for introverts and extroverts.
- **Do introverts and extroverts differ in their typical scores for your predominant color?**
- **Based on the MAD, which group (introverts or extroverts) has more variability for your predominant color's scores?**

## On your own

- Do introverts and extroverts in your class differ in their color scores? 
    - **Perform an analysis that produces _numerical summaries_ and _graphs_.**
    - **Then, write a few sentence interpretations that addresses this statistical question and considers the _shape_, _center_ and _spread_ of the distributions of the graphs you create.**
