---
title: "R's Normal Distribution Alphabet"
author: "Lab 2I"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Where we're headed

- In the last lab, you were able to overlay a normal curve on histograms of data to help you decide if the data's distribution is close to a normal distribution.
    - We also saw that calculating the `mean` of random shuffles also produces differences that are normally distributed.
- In this lab, we'll learn how to use some other `R` functions to:
    - Simulate random draws from a normal distribution.
    - Calculate probabilities with normal distributions.


## Get set up

- <span class="blue">Start by loading the `titanic` data and calculate the `mean` `age` of people in the data but `shuffle` their `survival` status 500 times.</span>
    - `Assign` this data the name `shfls`.
- <span class="blue">After creating `shfls`, use `mutate` to add a new variable to the dataset. This new variable should have the name `diff` and should be the `mean` `age` of those who survived minus those who died.</span>
- <span class="blue">Finally, calculate the `mean` and `sd` of the `diff` variable.</span>
    - `Assign` these values the name `diff_mean` and `diff_sd`.


## Is it normal?

- Before we proceed, we need to verify that our `diff` variable looks approximately normally distributed.
    - **Is the distribution close to normal? Explain how you determined this. Describe the center and spread of the distribution.**
    - **Compute and write down the mean difference in the age of the _actual_ survivors and the _actual_ non-survivors.**


## Using the normal model

- Since the distribution of our `diff` variable appears normally distributed, we can use a normal model to estimate the probability of seeing differences that are more extreme than our actual data.

- **Draw a sketch of a normal curve. Label the mean age difference, based on your shuffles, and the actual age difference of survivors minus non-survivors from the actual data. Then shade in the areas, under normal the curve, that are _smaller_ than the actual difference.**
    
- <span class="blue">Fill in the blanks to calculate the probability of an even smaller difference occurring than our actual difference using a normal model.</span>
```{r, eval = FALSE, echo = TRUE}
pnorm(____, mean = diff_mean, sd = ____)
```


## Extreme probabilities

- The probability you calculated in the previous slide is an estimate for how often we expect to see a difference smaller than the actual one we observed, by chance alone.

- <span class="blue">If you wanted to instead calculate the probability that the difference would be larger than the one observed, we could run (fill in the blanks):</span>
```{r, eval = FALSE, echo = TRUE}
1 - pnorm(____, mean = diff_mean, sd = ____)
```


## Simulating normal draws

- We can simulate random draws from a normal distribution with the `rnorm` function.
- <span class="blue">Fill in the blanks in the following two lines of code to simulate 100 heights of randomly chosen men. Assume the `mean` height is 67 inches and the `standard deviation` is 3 inches. </span> 

```{r, eval = FALSE, echo = TRUE}
draws <- rnorm(____, mean = ____, sd = ____)
```

- <span class="blue">Plot your simulated heights with a `histogram`.</span>

```{r, eval = FALSE, echo = TRUE}
histogram(draws, fit = ____)
```


## P's and Q's

- We've seen that we can use `pnorm` to calculate _probabilities_ based on a specified _quantity_.
    - Hence, why we call it "P" norm.
- Now we'll see how to do the opposite. That is, calculate a the _quantity_ for a specific _probability_.
    - Hence why we'll call this a "Q" norm.
- **How tall can a man be and still be in the shortest 25% of heights if the mean height is 67 inches with a standard deviation of 3 inches?**

```{r, eval = FALSE, echo = TRUE}
qnorm(____, mean = ____, sd = ____)
```

## On your own

Conduct one of the statistical investigations below:

- Using the `titanic` data:
    - **Were women on the Titanic typically younger than men?**
    - **Use a histogram, 500 random shuffles and a normal model to answer the question in the bullet above.**
    
    
- Using the `cdc` data:
    - **Using 500 random shuffles and a normal model, how much taller would the typical male have to be than the typical female in order for the difference to be in the upper 1% by chance alone?**
    - **How can we use this value to justify the claim that the average `Male` in our data is taller than the average `Female`?**
    
    
    
