---
title: "Text Analysis"
author: "The field of _Natural Language Processing_ has become EXTREMELY popular in recent years. Companies want to gain insights about people/topics based on written words. Writing software which analyzes the meanings of words (in context) is VERY difficult but can also be VERY valuable to companies, researchers, and data scientists."
date: "Outcome: You will perform a text analysis of the novel _Pride & Prejudice_ by Jane Austen"
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# library(mobilizr)
```


## Expanding **R**  

One reason **R** is popular amongst data scientists is because they are able to expand the functionality of **R** by installing _packages_.

_Packages_ are new sets of functions that **R** can use to perform tasks.

Cutting edge data science methods can be developed and shared with other **R** users.

New functions which make difficult programming tasks easier can be developed and shared via packages.  

### <b>Installing and loading packages</b>  

To install a new package (normally only performed once) use the command below:

```{r, eval = FALSE,echo=TRUE}
install.packages("package_name")
```

After installing a package, the new functions can be <b>loaded</b> into **R** for use by using the `library()` function

```{r, eval = FALSE, echo=TRUE}
library(package_name)
```

You should only ever install a new package exactly once. Even if you close **R**, the packages will remain _installed_.
Packages should be <b>loaded</b> with `library()` each time you restart **R**.  

## Packages Needed for Text Analysis

To help us begin to analyze words/text, the `tidytext` package needs to be installed.  

Before installing `tidytext`, click on the _Packages_ tab in RStudio (next to the Plots tab).  

Notice that the packages that are installed appear in alphabetical order.  Scroll down to verify if `tidytext` is already installed.  

```{r, eval = FALSE}
# This might take a minute
install.packages("tidytext")
```

If `tidytext` has been installed, load the functions contained in the package by:  

1. Checking off the `tidytext` box in the _Packages_ tab.

2. Using the `library()` command.  

```{r, warning = FALSE}
library(tidytext)
```

In this lab, we'll analyze the book _Pride and Prejudice_ by Jane Austen.  

Load the `janeaustenr` package (Which was installed when you installed `tidytext`) and then run
```{r, echo = FALSE}
library(janeaustenr)
```

```{r,echo=TRUE}
data(prideprejudice)
```

**What happens when you type the code below?**

```{r, eval = FALSE,echo=TRUE}
prideprejudice
```


### When installing packages goes bad

Install the `wordcloud` package if it does not appear in the _Packages_ tab.  

Load the functions contained in the package.


```{r, eval = FALSE}
install.packages("wordcloud")
```


```{r, eval = FALSE}
library(wordcloud)
```


### Installing older versions

You'll sometimes find that you need older versions of packages. Often times this is because packages rely on other packages that only run on the newest versions of **R**.

To install an older package version, first run

```{r, eval = FALSE,echo=TRUE}
library(devtools)
```

Now we can install `wordcloud` version 1.4

```{r, eval = FALSE,echo=TRUE}
install_version("wordcloud", version = 1.4)
```

Once you're done installing `wordcloud`, load the functions into **R**.

## Cleaning your text data

- Now that our packages are installed and loaded, it's time to get to work.
- Our first task is to turn the `prideprejudice` text into a `data.frame`.
- To do this, run:

```{r, eval = FALSE,echo=TRUE}
pride <- data.frame(line = 1:length(prideprejudice),
                    text = prideprejudice)
```

- Now, use the `mutate()` and `as.character()` functions to change the `text` variable from a `factor` to a `character`

### More cleaning

Our next goal is to break up each line of text into individual words. 

```{r, eval = FALSE, echo=TRUE}
pride <- unnest_tokens(pride, word, text)
```

`View` this `pride` data and note the new format.  

- NOTE: If you are analyzing text that is a variable in a data frame, use the command 
```{r,eval=FALSE,echo=TRUE}
select(data,variable)
```
to create a subset of the data frame that contains the single text variable.


## Basic Wordcloud

### Counting words

We're interested in seeing which words were used most often by Jane Austen, so we want to count the number of times each word appears in the book. Use `tally()` to accomplish this and assign the output the name `pride_words`.

- NOTE: Include the argument `format = "data.frame"` so the word counts become a new data frame.

### Plotting words

One very popular method of displaying text data is with word clouds. Create a wordcloud for the most 100 used words by running.

```{r, eval = FALSE, echo=TRUE}
with(pride_words, wordcloud(word, Freq, max.words = 100))
```

- **Notice any problems?**

### Stop words

When analyzing text, we're not often interested in common words like "I", "a", "the". We call these sort of _boring_ words _stop words_ (because we should _stop_ using them in our analysis).

`tidytext` helpfully comes equipped with a set of stop words we can use. To do so, run

```{r, eval = FALSE, echo=TRUE}
pride_words <- anti_join(pride_words, stop_words)
```

An `anti_join` is an efficient way to filter out words in `pride_words` that are also in `stop_words`.

- Recreate your wordcloud.  
**Does it look more useful?  
Which words were the most frequent?  
What happens if you include the argument `color = "steelblue"` within the `wordcloud` function?**

## Using sentiments

Words carry meaning. And we might be interested in altering our plot to show these different meanings. 

Run the following code to:  

(1) get a set of words with _positive_ and _negative sentiments (according to Bing) and  

(2) keep only the words from _Pride & Prejudice_ that contain sentiments

```{r, eval = FALSE, echo=TRUE}
sentiments <- get_sentiments(lexicon = "bing")
pride_sentiments <- inner_join(pride_words, sentiments)
```

- An `inner_join` keeps just the words in `pride_words` that also appear in `sentiments`

### Color & Sentiment

Now that we have sentiments assigned to our words, our next goal is to assign colors to each word based on its sentiment.  

To do this, we `mutate` our `pride_sentiments` and assign colors with the `ifelse()` function.

```{r, eval = FALSE, echo=TRUE}
pride_sentiments <- mutate(pride_sentiments,
                           colors = ifelse(sentiment == "positive", 
                                          yes = "steelblue", 
                                          no = "darkred"))
```

Re-create your word cloud one more time. This time though include the argument `color = colors` in your wordcloud function.

**1. Which positive sentiment words appeared most often?  
2. Which negative sentiment words appeared most often?  
3. Are there any words you feel were miscategorized?**

  