---
title: "Get the picture?"
author: "Lab 1B"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
---

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
data(cdc)
```

## Where'd we leave off ...

- In the previous lab, we started to get acquainted with the layout of RStudio and some of the commands.
- In this lab, we'll learn about different _types_ of variables. 
    - Such as those that are measured by numbers and others that have values that are categories.
- We'll also look at ways to visualize these different types of data using _plots_ (A word data scientists use interchangeably with the word _graph_).
- Find the _History_ tab in RStudio and click on it. Figure out how to use the information to reload the `cdc` `data`.


## Variable Types

- Numerical variables have values that are measured in units.
- Categorical Variables have values that describe or categorize our observations.
- `View` your `cdc` data and find the columns for `height` and `gender` (Use the _History_ pane again if you need help to `View` your data).
    - **Is `height` a numerical or categorical variable? Why?**
    - **Is `gender` a numerical or categorical variable? Why?**
    - **List either the different categories or what you think the measured units are for `height` and `gender`.**

    
## Which is which?

- Run the code you used in the previous lab to display the `names` of your `cdc` data's variables (Use the code displayed in the _History_ pane to resubmit previously typed commands). Use the code's output to help you complete the following:
    - **Write down 3 variables that you think are _categorical_ variables and why.**
    - **Write down 3 variables that you think are _numerical_ variables and why.**


## Data Structures

- One way to get a good summary of your data is to look at the data's _structure_.
    - One way to view this info would be to click on the little blue arrow next to `cdc` in the _Environment_ pane.
    - Another way would be to run the following in the console:
```{r, eval = FALSE, echo = TRUE}
str(cdc)
```

- Look at the `str`ucture of your `cdc` data and answer:
- **List all the types of info the `str()` function outputs**
- **Were you able to correctly guess which variables were categorical and numeric? Which ones did you mis-label?**
    

## Visualizing data

- Visualizing data is a really helpful way to learn about our variables.
    - Making a picture of the distribution of a variable is a good way to begin visualizing data.
    - Remember: A distribution gives us the values of the variable and tells us how many of these values we have in our data set.
- Choose one numeric and one categorical variable from the data and create both a `bargraph` and a `histogram` for each variable.
    - **Which function, either `bargraph` or `histogram` is better at visualizing categorical variables? Which is better at visualizing numerical variables?**
    

## How often do people text & drive?

- Make a graph that shows how often people in our data texted while driving.
    - **What does the y-axis represent?**
    - **What does the x-axis tell us?**
    - **Would you say that _most_ people _never_ texted while driving? What does the word _most_ mean?**
    - **Approximately what percent of the people texted while driving for 20 or more days? (Hint: There's `r nrow(cdc)` students in our data.**


## Does texting and driving differ by gender?

- Fill in the blanks with the correct variables to create a side-by-side bargraph:
<div class='greenequation'> _bargraph_ (~ ____ , data = ____ , groups = ____ ) </div>
- **Write a sentence explaining how boys and girls differ in their use texting while driving.**
- **Would you say that most girls never text and drive?  Would you say that most boys never text and drive?**
- **How did including the `groups` argument in your code change the graph?**


## Do males/females have similar heights?

- To answer this, what we'd like to do is visualize the distributions of heights, separately, for males and females.
    - This way, we can easily compare them.
- Use the `groups` argument to create a `histogram` for the `height` of males and females.
    - **Can you use this graphic to answer the question at the top of the slide? Why or why not?**
    - **Is grouping numeric values, such as heights, as helpful as grouping categorical variables, such as texting & driving?**
    

## Do males/females have similar heights?

- `groups` uses color to differntiate between groups.
    - **Why does this work for bargraphs but not for for histograms?**
- Fill in the blanks with the correct variables to create a split histogram (The " | " symbol is usually between the delete and enter keys on a keyboard) to answer the questions below:
<div class='greenequation'> _histogram_ (~ ____ | ____ , data = ____ ) </div>
- **Do you think males & females have similar heights? Use the plot you create to justify your answer.**
- **Just like we did for the histogram, is it possible to create a _split_ bargraph? Try to create a bargraph of `drive_text` that's split by `gender` to find out.**


## On your own:

- In this lab, we looked at boy's and girl's attitudes about using texting & driving: 
- **What other factors do you think might affect how often people text and drive?**
    - **Choose one variable from the `cdc` data, make a graph, and use the graph to describe how `drive_text` use differs with this variable.**