---
title: "Visualizations with ggplot"
author: "Advanced PD Lab 2B"
date: The following slides contain information and examples about how to create visualizations
  using ggplot. You will need your ggplot CHEAT SHEET.
output:
  ioslides_presentation: default
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
mpg<- mutate(mpg,cyl=as.factor(cyl))
```

## What is ggplot?

The ggplot package, created by <a href="http://hadley.nz/">Hadley Wickham</a>, Chief Scientist at RStudio, offers a powerful graphics language for creating elegant and complex plots.  

ggplot allows you to create graphs that represent both univariate and multivariate numerical and categorical data in a straightforward manner.  

**Let's get started:** 

Begin by installing and loading the `ggplot` package.  

`install.packages(" ")`  

`library( )`

### Consider Data:  

Load and preview the `mpg` data. 

 
## Categorical Variables  

The template to create a visual in ggplot looks like:  



    ggplot(data = < DATA >) +  

    < GEOM_FUNCTION >(mapping = aes(< MAPPINGS >))  
    
    


**Include all of the codes you come across in an RMarkdown:**  

**Cheat Sheet**  

<b>ONE VARIABLE discrete</b>

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class))
```



The following arguments can be applied to change the color of a visual: `color, fill`  



```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class), fill="steelblue")
```

**1. Replace the `fill` argument with `color` and explain how it affects the visual.**  


**2. What code will create the visual below?** 


```{r, eval=TRUE, echo=FALSE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class), fill="white",color="black")
```

The `fill` argument can also be used to assign colors to each category of a categorical variable (similar to what the groups argument does in lattice).  Notice that the argument needs to be contained within the aesthetics.  

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill=class))
```


Let's add more layers to this visual. 

**CHEAT SHEET**  

<b>Scales, Labels, Coordinate Systems and Themes</b>

Functions from Scales, Labels, Coordinate Systems and Themes have been applied to modify the visual.  

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = class)) +
  scale_y_continuous(breaks = seq(0, 70, 5))+
  labs(title ="Distribution of Class") +
  coord_flip ()+
  theme_classic ()
```

**Flex Your Skills:** 

**3. Modify the visual above by changing the scale, including additional labels and changing the theme.**  


## Comparing Groups on a Categorical Variable

<b> Is there a difference in the number of cylinders between types of cars?</b>

In order to create a visual to compare groups using lattice, we utilized the vertical bar (`|`) to split or facet each groups' data into a separate graph. The `groups` argument color coded the groups.

ggplot offers similar options with their `Faceting` and `Position Adjustment` options.  

<B>NOTE:</b> If you look at the structure of the mpg data you will find that the variable `cyl`is being interpreted as an integer. We, however, are utilizing it more as a categorical variable so we need to mutate the data using the following code:

```{r, echo=TRUE}
mpg<-mutate(mpg, cyl=as.factor(cyl))
```

Let's look at the most basic code in ggplot2 for creating a bargraph comparinggroups...in this case the number of cylinders between type of car.

```{r}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class,fill=cyl))
```

4. **What do you predict will happen if you swap `class` and `cyl` in the code above? Check if your prediction is correct.**

Let's see the different ways we can facet the visuals!

**Cheat Sheet**  

<b>Faceting</b> 

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = cyl,fill=cyl))+
  facet_grid(cols=vars(class))
```

**5. What code will create the visual below?** 

```{r, eval=TRUE, echo=FALSE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = cyl,fill=cyl))+
  facet_grid(rows=vars(class))
```

**6. What code will create the visual below?** 

```{r, eval=TRUE, echo=FALSE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = cyl,fill=cyl))+
  facet_wrap(vars(class))
```

Let's see what the position functions do!

**Cheat Sheet**  

<b>Position Adjustments</b>  

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class,fill=cyl),position="dodge")
```

**7. What do you predict will happne if you swap `class` and `cyl` in the code above? Check and see if your prediction was correct.**   

Let's look at the `fill` position.

```{r, eval=TRUE, echo=TRUE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class,fill=cyl), position="fill")
```

**8. When might we want to use the `fill` position? ** 


## Numerical Variables 

**CHEAT SHEET**  

ONE VARIABLE continuous</b>

Let's look at a basic histogram.

```{r, echo=TRUE,eval=TRUE}
ggplot(data=mpg)+
  geom_histogram(mapping=aes(x=cty))
```

Notice that a histogram was created but the computer was not happy about the number of bins (30). This histogram is behaving more like a dotplot because there is a bin for each value of city miles per gallon. The `binwidth` argument works similar to the `nint` argument in lattice. Currenlty there is one bin for each value (binwidth=1). 

```{r, echo=TRUE,eval=TRUE}
ggplot(data=mpg)+
  geom_histogram(mapping=aes(x=cty),binwidth = 3)
```

**9. What code will create the visual below?** 


```{r, echo=FALSE,eval=TRUE}
ggplot(data=mpg)+
  geom_histogram(mapping=aes(x=cty),binwidth = 2,fill="orange")+
  labs(x="city miles per gallon")+
  theme_linedraw()
```

ggplot allows us to overlay plots!


```{r, eval=TRUE, echo=TRUE}
ggplot(data=mpg)+
  geom_histogram(mapping=aes(x=cty),binwidth = 2)+
  geom_dotplot(mapping=aes(x=cty),fill="white",binwidth = 0.6)
```

**10. Change the order of the plots for the code above and explain what happens?** 

```{r, eval=FALSE, echo=FALSE}
ggplot(data=mpg)+
  geom_dotplot(mapping=aes(x=cty),fill="white",binwidth = 0.6)+
  geom_histogram(mapping=aes(x=cty),binwidth = 2)
  
```

**11. What code will create the visual below?**  

```{r, eval=TRUE, echo=FALSE}
ggplot(data=mpg)+
  geom_dotplot(mapping=aes(x=cty),binwidth=0.2)+
  facet_grid(rows=vars(cyl))
```


## TWO VARIABLES  

**CHEAT SHEET**  

<b>discrete x, continuous y</b>

Notice that the geom_boxplot() function calls for a discrete variable
and a continuous variable. The geom_boxplot() function is BEST suited to 
compare groups on a numerical variable.


```{r, eval=TRUE, echo=TRUE}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x=class,y=hwy,color=class))
  scale_y_continuous(breaks=seq(0,45,5))
```


**12. Interchange the x and y variable. Explain what happens.**

We can trick the geom_boxplot() function into plotting a single variable by 
changing the numerical variable into a character.

```{r, eval=TRUE, echo=TRUE}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x="hwy",y=hwy))
```

**13. What code will create the visual below?**  

```{r, eval=TRUE, echo=FALSE}
ggplot(data=mpg)+
  geom_boxplot(mapping=aes(x="cty",y=cty))+
  coord_flip()+
  scale_y_continuous(breaks=seq(10,36,1))+
  theme_bw()
```


**CHEAT SHEET**  

<b>continuous x, continuous y</b>  


<b>Is there a difference between city miles per gallon and highway miles per gallon?</b>  
In order to help us answer this question we need to add the equation y = x to our scatterplot. The line y = x represents when hwy mileage = cty mileage (no difference).  

```{r, eval=TRUE, echo=TRUE }
ggplot(data=mpg)+
  geom_point(mapping=aes(x=cty,y=hwy))+
  geom_abline(intercept = 0, slope = 1, color="red")
```

**14. What code will create a scatterplot of hwy vs cty with the line of best fit?**

```{r, eval=FALSE, echo=FALSE }
ggplot(data=mpg)+
  geom_point(mapping=aes(x=cty,y=hwy))+
  geom_abline(intercept = 0.892, slope =1.337 , color="blue")
```


<b>What is the relationship between highway mileage and displacement?</b>  


```{r, eval=TRUE, echo=FALSE}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy))
```

Notice that as engine displacement increases, highway miles per gallon decreases. There are a few observations, however, that don't quite fit the pattern.  

For those of you that want to learn about engine displacement, watch this short video: <iframe frameborder="0" scrolling="no" marginheight="0" marginwidth="0"width="788.54" height="443" type="text/html" src="https://www.youtube.com/embed/bf1dOSPB2UA?autoplay=0&fs=0&iv_load_policy=3&showinfo=0&rel=0&cc_load_policy=0&start=0&end=0&origin=https://youtubeembedcode.com"><div><small><a href="https://youtubeembedcode.com/nl/">youtubeembedcode nl</a></small></div><div><small><a href="http://tr3ndygirl.com/">Ultimate web Traffic</a></small></div></iframe>

Try the codes below: determine which variable explains these outliers.  

```{r, eval=FALSE, echo=TRUE}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,shape=cyl))
```

```{r, eval=FALSE, echo=TRUE}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,size=year))
```

```{r, eval=FALSE, echo=TRUE}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,color=fl))
```

```{r, eval=FALSE, echo=TRUE}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy,color=class))
```

**15. What are the different aesthetics that can be applied to a geometric object (point)?**

## On Your Own  

There are many more non-traditional plots that can be made using ggplot. 

A 2d density chart displays the relationship between 2 numeric variables. One is represented on the X axis, the other on the Y axis, like for a scatterplot. Then, the number of observations within a particular area of the 2D space is counted and represented by a color gradient. Several types of 2d density charts exist.

**CHEAT SHEET**  

<b>continuous bivariate distribution</b> 

```{r, eval=TRUE, echo=FALSE}
ggplot(diamonds,aes(carat,price))+
  geom_density2d()
  
```

In order to create this graph notice on the cheat sheet it has `h + geom_density2d()`  

Notice that `h` refers to `ggplot(diamonds,aes(carat,price))`

So the graph above was created using the following code:  

```{r, eval=FALSE, echo=TRUE}
ggplot(diamonds,aes(carat,price))+
  geom_density2d()
  
```

This is equivalent to:

```{r, eval=FALSE, echo=TRUE}
ggplot(data=diamonds)+
  geom_density2d(mapping=aes(x=carat,y=price))
```

**16. Chose a different visualization from the ggplot cheat sheet to create.** 