---
title: "Finding clusters"
author: "Lab 4H"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Clustering data

- We've seen previously that data scientists have methods to predict values of specific variables.
    - We used _regression_ to predict numerical values and _classification_ to predict categories.
- _Clustering_ is similar to classification in that we want to group people into categories. But there's one important difference:
    - In _clustering_, we don't know how many groups to use because we're not predicting the value of a known variable!
- In this lab, we'll learn how to use the k-means clustering algorithm to group our data into clusters.


## The k-means algorithm

- The k-means algorithm works by splitting our data into _k_ different clusters.
    - The number of clusters, the value of _k_, is chosen by the data scientist.
- The algorithm works _only_ for numerical variables and _only_ when we have no missing data.
- To start, use the `data` function to load the `futbol` data set.
    - This data contains 23 players from the US Men's National Soccer team (USMNT) and 22 quarterbacks from the National Football League (NFL).
- Create a scatterplot of the players `ht_inches` and `wt_lbs` and color each dot based on the `league` they play for.


## Running k-means

- After plotting the player's heights and weights, we can see that there are two clusters, or different types, of players:
    - Players in the NFL tend to be taller and weigh more than the shorter and lighter USMNT players.
- Fill in the blanks below to use k-means to cluster the same height and weight data into two groups:

```{r, eval = FALSE, echo = TRUE}
kclusters(____~____, data = futbol, k = ____)
```

- Use this code and the `mutate` function to add the values from `kclusters` to the `futbol` data. Call the variable `clusters`.


## k-means vs. ground-truth

- In comparing our football and soccer players, we _know_ for certain which league each player plays in.
    - We call this knowledge _ground-truth_.
- Knowing the _ground-truth_ for this example is helpful to illustrate how k-means works, but in reality, data-scientists would run k-means not knowing the _ground-truth_.
- **Compare the clusters chosen by k-means to the ground-truth. How successful was k-means at recovering the `league` information?**


## On your own

- Load your class' `timeuse` data (remember to run `timeuse_format` so each row represents the mean time each student in spent participating in the various activities).
- Create a scatterplot of `homework` and `videogames` variables.
    - Based on this graph, identify and remove any outliers by using the `subset` function.
- Use `kclusters` with `k=2` for `homework` and `videogames`. 
    - **Describe how the groups differ from each other in terms of how long each group spends playing `videogames` and doing `homework`.**
