---
title: "Data, Code & RStudio"
author: "Lab 1A"
date: 'Directions: Follow along with the slides and answer the questions in **red**
  font in your journal.'
output:
  word_document:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
data(cdc)
```

## Welcome to the labs!

- Throughout the year, you'll be putting your data science skills to work by completing the labs.
- You'll learn how to program in the `R` programming language.
    - The programming language used by actual data scientists.
- Your code will be written in RStudio which is an easy to use interface for coding using `R`.


## So let's get started!

- The data for our first few labs comes from the Centers for Disease Control (CDC)
    - The CDC is a federal institution that studies public health.
- Type these two commands into your console:

```{r, eval = FALSE, echo = TRUE}
data(cdc)
```
```{r, eval = FALSE, echo = TRUE}
View(cdc)
```
  
- **Describe the data that appeared after running `View(cdc)`:** 
    - **_Who_ is the information about?**
    - **What sorts of information about them was collected?**

- To find out more information about the `cdc` data, type the command below into your console.
  - To get back to the slides find and click on the `Viewer` tab.

```{r, eval = FALSE, echo = TRUE}
?cdc
```



## Data: Variables & Observations

- Data can be broken up into two parts. 
    1. _Observations_
    2. _Variables_

  - _Observations_ are the _who_ or _what_ we are collecting data from/about.
  - _Variables_ are the measurements or characteristics about our _observations_.
- If need be, re-type the command you used to `View` your data. Then answer the following:
    - **Based on the data, describe a few characteristics about the first observation.**
    - **What does the first column tell us about our observations?**  
    
- In order to describe the first observation, notice that you had to look at the first row of the spreadsheet. Each row, in this case, describes a person.
- The columns of the spreadsheet represent variables.

## Uncovering our Data's Structure

- Now that we've looked at our data, let's look at how RStudio is organized.
- RStudio's main window is composed of four _panes_
- Find the pane that has a _tab_ titled _Environment_ and click on the _tab_.
    - This pane contains a list of everything that's currently available for R to use.
    - Notice that R knows we have our `cdc` data loaded.
- **How many students are in our `cdc` data set?**
- **How many variables were measured for each student?**

  
## Type the following commands into the console

```{r, echo=TRUE, results='hide'}
dim(cdc)
```
```{r, echo=TRUE, results='hide'}
nrow(cdc)
```
```{r, echo=TRUE, results='hide'}
ncol(cdc)
```
```{r, echo=TRUE, results='hide'}
names(cdc)
```

- **Which of these functions tell us the number of observations in our data?**
- **Which of these functions tell us the number of variables?**


## First Steps

- Typing commands into the console is your first step into the larger world of _programming_ or _coding_ (terms which are often used interchangeably).
- Coding is all about learning how to send instructions to your computer.
    - The way we _speak_ to the computer, using a coding language, is _syntax_.
- `R` is one of many coding languages. Each coding language is slightly different, and these differences are reflected in the syntax.
- _Capitalization_, _spelling_ and _punctuation_ are REALLY important.


## Syntax matters

- **Run the following commands and write down what happens after each. Which does R understand?**

```{r, eval = FALSE, echo = TRUE}
Names(cdc)
```
```{r, eval = FALSE, echo = TRUE}
NAMES(cdc)
```
```{r, eval = FALSE, echo = TRUE}
names(cdc)
```
```{r, eval = FALSE, echo = TRUE}
names(CDC)
```


## R's most important syntax

<div class='greenequation'> _function_ (y~x, data = ____ )  </div>

- Notice the command and its corresponding plot. 

```{r, echo=TRUE, eval=TRUE, fig.width=6, fig.height=4}
bargraph(~grade, data = cdc)
```

- Most of the commands you will be using follow the syntax below:

<div class='greenequation'> _function_ (y~x, data = ____ )  </div>

- To create graphs or plots you need to provide `R` with the following:
  - The name of the R function, often the plot’s name, that tells the computer how to create your graph.
  - The variable(s) containing the information we want the function to use.
  - The data set containing the variables.

- Later on, we’ll see we can use this syntax to do more than create graphs.

    

## Syntax in action

<div class='greenequation'> _function_ (y~x, data = ____ )  </div>

- Search through the different panes. Find and then click on the _Plots_ tab.
    - To get back to the slides, find and then click on the _Viewer_ tab.

- **Which one of these plots would be useful for answering the question: _Is it unusual for students in the CDC dataset to be taller than 1.8 meters?_**

- Run the three commands below then answer the question that follows.

```{r, echo=TRUE, eval=FALSE}
histogram(~height, data = cdc)
```
```{r, echo=TRUE, eval=FALSE}
bargraph(~drive_text, data = cdc)
```
```{r, echo=TRUE, eval=FALSE}
xyplot(weight~height, data = cdc)
```

- **Do you think it's unusual for students in the data to be taller than 1.8 meters? Why or why not?**

- Hint: Use the arrow keys on the `Plots` tab to toggle between the plots.

## On your own:

- After completing the lab, answer the following questions:
    - **What is _public health_ and do we collect data about it?**
    - **How do you think our data was collected? Does it include every high school aged student in the US?**
    - **How might the CDC use this data? Who else could benefit from using this data?**
    - **Write the code to visualize the distribution of weights of the students in the CDC data with a histogram. What is the _typical_ weight?**
    - **Write the code to create a bargraph to visualize the distribution of how often students ate fruit. About how many students did not eat fruit over the previous 7 days?**