---
title: "What's the FREQ?"
author: "Lab 1G"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(mobilizr)
data(atu_clean)
```

## Clean it up!

- In Lab 1F, we saw how we could _clean_ data to make it easier to use and analyze.
    - You cleaned a small set of variables from the American Time Use (ATU) survey.
    - The process of cleaning and then analyzing data is _very_ common in Data Science.
- In this lab, we'll learn how we can create frequency tables to detect relationships between categorical variables.
    - <span class="blue">For the sake of consistency, rather than using the data that you cleaned, you will use the pre-loaded ATU data.</span>
    - <span class="blue">Use the `data()` function to load the `atu_clean` data file to use in this lab.</span>



## How do we summarize categorical variables?

- When we're dealing with categorical variables, we can't just calculate an _average_ to describe a _typical_ value.
    - (Honestly, what's the average of categories _orange_, _apple_ and _banana_, for instance?)
- When trying to describe categorical variables with numbers, we calculate _frequency tables_.


## Frequency tables?

- When it comes to categories, about all you can do is _count_ or _tally_ how often each category comes up in the data.
- **Fill in the blanks below to answer the following: How many more _females_ than _males_ are there in our ATU data?**
```{r, eval = FALSE, echo = TRUE}
tally(~ ____, data = ____)
```


## 2-way Frequency Tables

- Counting the categories of a single variable is nice, but often times we want to make comparisons.
- For example, what if we wanted to answer the question:
  - Does one `gender` seem to have a higher occurrence of physical challenges than the other? If so, which one and explain your reasoning?
  
- We could use the following plot to try and answer this question:
  
```{r, eval=TRUE, echo=TRUE, fig.width=7, fig.height=5}
bargraph(~phys_challenge | gender, data = atu_clean)
```

- The split bargraph helps us get an idea of the answer to the question, but we need to provide precise values.
- <span class="blue">Use a line of code, that's similar to how we facet plots, to obtain a tally of the number of people with physical challenges and their genders.</span>
- **Write down the resulting table.**



## Interpreting 2-way frequency tables

- Recall that  there were `r tally(~gender, data = atu_clean)[2] - tally(~gender, data = atu_clean)[1]` more women than men in our data set.
    - If there are more women, then we might expect women to have more physical challenges (compared to men).
- Instead of using _counts_ we use _percentages_.
- Include: `format = "percent"` as an option to the code you used to make your 2-way frequency table. Now answer this question: 
    - **Does one `gender` seem to have a higher occurrence of physical challenges than the other? If so, which one and explain your reasoning.**

- It's often helpful to display totals in our 2-way frequency tables.
    - To include them, include `margins = TRUE` as an option in the tally function.
    
## Conditional Relative Frequencies

- There is as difference between `phys_challenge | gender` and `gender | phys_challenge`. 

```{r, eval=TRUE, echo=TRUE}
tally(~phys_challenge | gender, data = atu_clean, margin = TRUE)
```

```{r, eval=TRUE, echo=TRUE}
tally(~gender | phys_challenge, data = atu_clean, margin = TRUE)
```
- At first glance, the two-way frequency tables might look similar (especially when the `margin` option is excluded). Notice, however, that the totals are different.
- The totals are telling us that `R` calculates conditional frequencies by column!

- What does this mean?
  - In the first two-way frequency table the groups being compared are `Male` and `Female` on the distribution of physical challenges.
  - In the second two-way frequency table the groups being compared are the people with `No difficulty` and those that `Has difficulty` on the distribution of gender.
- **Add the option `format = "percent"` to the first tally function. How were the percents calculated? Interpret what they mean.**

## On your own

- **Describe what happens if you create a 2-way frequency table with a numerical variable and a categorical variable.**
- **How are the types of statistical questions that 2-way frequency tables can answer different than 1-way frequency tables?**
- **Which gender has a higher rate of _part time employment_?**

