---
title: "Have I Lost My Marbles?"
output: pdf_document
---

Directions: Follow along with the slides and answer the questions in **BOLDED** font in your journal.
```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(MobilizeSimple)
```

Some background...
===============================
- Let's assume we have a bag of 100 marbles.
  - There are 23 <font color = "blue"> blue </font> marbles, and the rest are <font color = "green"> green </font>.

  - **What is the probability of selecting one <font color = "blue"> blue </font> marble from this bag?**
  
- In this lab, we will be _estimating_ **probabilities** by drawing  marbles from the bag and recording our responses.

But wait! I didn't bring a bag of marbles...
===============================
- No problem! You can **simulate** one in RStudio!
- Let's first create a vector for just the blue marbles.

```{r,eval=FALSE}
blue <- rep("blue", times = 23)
```
  - **What do you think the `rep()` function does?**


Finishing up the bag
===============================
- Create a vector for the green marbles on your own and name it `green`. 
    - **Write down your command in your DS Journal.**
    
- Next, _combine_ the `blue` and `green` vectors into one new vector and name it `marbles`.
  - **Write down your command in your DS Journal.**

Now we can create our first sample!
===============================
- If we just want to draw ONE marble out, we can run the following code:

```{r,eval=FALSE}
sample1 <- sample(x = marbles, size = 1)
```

- **What color marble did you draw?**

- **If we wanted a sample of 10 marbles instead of just one, how could we revise the code?**

- Run your revised code and **write down the sample of 10 marbles in your DS Journal.**

But wait!
================================
- We can select marbles one after the other, or put the marble back into the bag each time we draw (like we did in the previous class).

- Within the `sample()` function, there is an option called `replace`, which we can set to either `TRUE` or `FALSE`.


To replace or not to replace
================================
- Let's take a sample of 101 marbles. That is, we'll draw a marble from our bag of 100, note its color, and then draw another. We will do this 101 times.
- Use the `replace` option - first set it equal to `FALSE`, and then try it with `TRUE`.

```{r,eval=FALSE}
sample(x = marbles, size = 101, 
       replace = FALSE)
```
- **What happens when `replace = FALSE`? Why do you think this happened?**
- **What happens when `replace = TRUE`? Which is better for this scenario?**


Back to those 10 marbles...
================================
- Rerun your code for the sample size of 10 (don't forget to create a new name for this sample - maybe `sample2`?) and include the `replace` argument. **Write down the new sample in your DS Journal.**
  
- **What percent of the 10 marbles were blue?**

$$\text{Estimated Probability} = \frac{\text{\# of Blue Marbles}}{\text{\# of Total Draws}}$$

- **How does your answer compare to your neighbor's?**


Sample probability vs. theoretical probability
=================================
- **How does your answer from the previous slide compare to the theoretical probability of selecting a blue marble?** (Refer back to your answer from Slide 2 if you forgot the true probability.)

- Let's run our code a few more times, but instead of just drawing 10 marbles from our bag, let's draw out 300.

- Obviously, we don't want to write out all 300 marble colors, so we can simply tally up the number of blue marbles that our sample produced by using the `tally()` function.


Sample probability vs. theoretical probability
=================================
- **In your sample of 300, how many marbles were blue? What is your estimated probability of selecting a blue marble?**

- **As a class, report your estimated probabilities. Examine them and make a guess about which value was the _typical_ estimated probability.**

  - Note: The estimated probabilities are all close to the theoretical probability.
  
  - The larger the sample size, `n`, the better our estimated probability will be.
  

For future study...
=================================
- **How would you design a simulation to find the probability that a second marble is blue?**
