---
title: "Which song plays next?"
author: "Lab 2C"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
rap <- rep("rap", times = 39)
rock <- rep("rock", times = 61)
songs <- c(rap, rock)
```

## A new direction

- For the past two labs, we've looked at ways that we can summarize data with numbers.
    - Specifically, you learned how to describe the _center_, _shape_ and _spread_ of variables in our data.
- In this lab, we're going to _estimate the probability_ that a rap song will be chosen from a playlist with both rap and rock songs, if the choice is made at random.
    - The playlist we'll work with has 100 songs: 39 are rap and 61 are rock.


## Estimate what ... ?

- To _estimate the probability_, we're going to imagine that we select a song at random, write down its genre (_rock_ or _rap_), put the song back in the playlist, and repeat 499 more times for a total of 500 times.
- The statistical question we want to address is: _On average, what proportion of our selections will be rap?_
- **Why do we _put a song back_ each time we make a selection?** 
- **What would happen in our little experiment if we did not do this?**  


## Calculating probabilities

- Remember that a _probability_ is the long-run proportion of time an event occurs.
    - Many probabilities can be answered exactly with just a little math. 
    - The probability we draw a single rap song from our playlist of 39 rap and 61 rock songs is `39/100`, `0.39` or `39%`.
- Probabilities can also be answered exactly if we were willing to randomly select a song from the playlist, write down its _genre_, place the song back in the list, and repeatedly do this _forever_.
    - Literally, _forever_ ...
    - But we don't have that much time. So we're only going to do it 500 times which will give us an _estimate of the probability_.


## Estimating probabilities

- You might ask, _Why are we estimating the probability if we know the answer is 39%?_  
    - Sometimes, probabilities are too hard to calculate with simple division as we did above. In which case, we can often program a computer to run an experiment to estimate the probability.
    - We refer to these programs as _simulations_.
- The techniques you learn in this lab could be applied to very simple probability calculations or very hard and complex calculations.
    - In both cases, your _estimated_ probability would be very close to the _actual_ probability.


## Getting ready

- Simulations are meant to mimic what happens in real-life using randomness and computers.
    - Before we can start simulating picking songs from a playlist, we need to simulate that playlist in `R`.
- <span class="blue">Simulate our 39 `rap` songs using the repeat `rep()` function.</span>
```{r, eval = FALSE, echo = TRUE}
rap <- rep("rap", times = 39)
```
- <span class="blue">Look in the `Environment` pane for the vector containing your `rap` songs.</span>
- <span class="blue">Use a similar line of code to simulate the rock songs in our playlist of 100.</span>


## Put the songs in the playlist

- Now that we've got some different songs, we need to combine them together.
    - To do this, we can use the combine function `c()` in `R`.
- <span class="blue">Fill in the blanks to combine your different songs:</span>
```{r, eval = FALSE, echo = TRUE}
songs <- __(rap, ____)
```
- And with that, our playlist of songs should be ready to go. 
    - <span class="blue">Type `songs` into the console and hit enter to see your individual _songs_.</span>


## Pick a song, any song

- Data scientists call the act of choosing things randomly from a set _sampling_.
    - We can randomly choose a song from our playlist by using:
```{r, eval = FALSE, echo = TRUE}
sample(songs, size = 1, replace = TRUE)
```

- <span class="blue">Run this code 10 times and compute the _proportion_ of `"rap"` songs you drew from the 10.</span>
  - Vocabulary Check: A <b>proportion</b> is a fraction of the whole.
    - For example, if 2 rap songs were drawn from the 10, the _proportion_ would be 2/10 
    - It is more common to express a _proportion_ as a decimal, in this case, 0.20
    - It is even more common to express a _proportion_ as a percentage,  20%

- **Once everyone in your class has computed their _proportions_, calculate the _range_ of _proportions_ (the largest _proportion_ minus the smallest _proportion_) for your class and write it down.**


## Now do() it some more

- Instead of running the same line of code multiple times ourselves we can use `R` to `do()` multiple repetitions for us.
    - <span class="blue">Fill in the blanks below to `do` the `sample` code from the previous slide _50_ times:</span>
```{r, eval = FALSE, echo = TRUE}
do(___) * sample(___, ___ = ___, ___ = ___)
```
- Recall that we need to store our results to be able to perform analysis.
- <span class="blue">_Assign_ the 50 selected songs the name `draws` and then `View` your file.</span>

- **What is the variable name?**
  - `R` defaulted to naming the variable based on the function used. You may use the data cleaning skills you learned in lab 6 to `rename` the variable if you wish.

- <span class="blue">Fill in the blank below to `tally` how often each genre was selected:</span>
```{r, eval = FALSE, echo = TRUE}
tally(~___, data = draws)
```
- **Compute the proportion of `"rap"` songs for your 50 draws and find out if the _range_ for your class' proportions is bigger or smaller than when we drew 10 songs.**


## Proportions vs. Probability

- To review, so far in this lab we've:
    - Simulated a "playlist" of songs.
    - Repeatedly simulated drawing a song from the playlist, noting its genre and placing it back in the playlist.
    - Computed the proportion of the draws that were `"rap"`.
- These proportions are all _estimates_ of the theoretical probability of choosing a rap song from a playlist.
    - As we increase the number of draws, the _range_ of proportions should shrink.

_When using simulations to estimate probabilities, using a large number of repeats is better because the estimates have less variability and so we can be confident we're closer to the actual value._


## Non-random Randomness

- We've seen that random simulations can produce many different outcomes.
    - Some estimated probabilities in your class were smaller/larger relative to others.
- There are instances where you might like the same random events to occur for everyone.
    - We can do this by using `set.seed()`.
- For example, the output of this code will always be the same:
```{r, eval = TRUE, echo = TRUE}
set.seed(123)
sample(songs, size = 1, replace = TRUE)
```


## Playing with seeds

- <span class="blue">With a partner, choose a number to include in `set.seed` then redo the simulation of 50 songs. </span>
    - Both partners should run `set.seed(___)` just before simulating the 50 draws. 
    - The blank in `set.seed(___)` should be the same number for both partners.
    - **Verify that both partners compute the same proportion of `"rap"` songs.**
- <span class="blue">Redo the 50 simulations one last time but have each partner choose a different number for `set.seed(___)`. </span>
    - **Are the proportions still the same? If so, can you find two different values for `set.seed` that give different answers?**


## On your own

- Suppose there are 1,200 students at your school. 400 of them went to the movies last Friday, 600 went to the park and the rest read at home.

_If we select a student at random, what is the probability that this student is one of the one's who went to the movies last Friday?_

- <span class="blue">Answer this by estimating the probability that a randomly chosen student went to the movies using 500 simulations.</span>
    - **Write down both the estimated probability and the code you used to compute your estimate. You might find it helpful to write your answer in an R Script _(File -> New File -> R Script)_**
    - <span class="blue">Include `set.seed(123)` in your code before you `do` 500 repeated samples.</span>