---
title: "Maps"
author: "Lab 3F"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Informative and Fun!

- Maps are some of the most interesting plots to make because the info represents:
  - Where we live.
  - Where we go.
  - Places that interest us.
- Maps are also helpful to display geographic information.
  - John Snow (the physician, not the character from _Game of Thrones_...) once famously used [a map to discover how cholera was transmitted.](http://commons.wikimedia.org/wiki/File:Snow-cholera-map-1.jpg#mediaviewer/File:Snow-cholera-map-1.jpg)
- In this lab, we'll use `R` to create an interactive map of the `mtns` data we scraped in Lab 3E.


## Getting ready to map

- The map we'll be creating will end up in RStudio's _Viewer_ pane.
    - Which means you'll need to alternate between building the map and loading the lab.
- You'll find it very helpful, for this lab, to write all of the commands, including the `load_lab(23)` command, as an `R` script.
    - This way you can edit the code that builds the map and quickly reload the lab.

## Load your data!

- In Lab 3E you created a dataset. Load it into Rstudio now by filling in the blank with the file name of the data.
```{r, eval = FALSE, echo = TRUE}
load("___.Rda")
```
- Didn't finish the lab or save the data file? Ask a friend to share it!


## Build a Basic Map

- Let's start by building a basic map!
- Use the `leaflet()` function and the `mtns` data to create the `leaf` that we can use for mapping.

```{r, eval = FALSE, echo = TRUE}
mtns_leaf <- leaflet(____)
```

- Then, insert `mtns_leaf` into the `addTiles()` function and assign the output the name `mtns_map`
- Run `mtns_map` in the console to look at your basic map with no data displayed.
    - Be sure to try clicking on the map to pan and zoom.


## Including our data

- Now we can add markers for the locations of the mountains using the `addMarkers()` function.
    - Fill in the blanks below with the basic map we've created and the values for latitude and longitude.
    
```{r, eval = FALSE, echo = TRUE}
addMarkers(map = ____, lng = ~____, lat = ~____)
```

- Supply the `peak` variable, in a similar way as we supplied the `lat` and `long` variables, to the `popup` argument and include it in the code above.
    - **Click on a marker within California and write down the name of the mountain you clicked on.**


## Colorize

- Our current map looks pretty good, but what if we wanted to add some colors to our plot? 
- Fill in the blanks below to create a new variable that assigns a color to each mountain based on the `state` its located.
```{r, eval = FALSE, echo = TRUE}
mtns <- mutate(____, state_colors = colorize(____))
```
- Now that we've added a new variable, we need to re-build `mtns_leaf` and `mtns_map` to use it.
    - Create `mtns_leaf` and `mtns_map` as you did before.
    - Then change `addMarkers` to `addCircleMarkers` and keep all of the arguments the same.


## Showing off our colors

- To add the colors to our plot, use the `addCircleMarkers` like before but this time include `color = ~state_colors` as an argument.
- It's hard to know just what the different colors mean so let's add a legend.
  - First, assign the map with the circle markers as `mtns_map`.
  - Then, fill in the blanks below to place a legend in the top-right hand corner.
```{r, eval = FALSE, echo = TRUE}
addLegend(____, colors = ~unique(____), labels = ~unique(____))
```
