---
title: "Lab Style Guide"
author: "mobilize-james"
date: "July 15, 2014"
output:
  html_document:
    css: IDSLabCSS.css
---

### Highlighting

#### **Bold**

- **Questions that students should answer in their Data Science Journals should be bolded.**
- **Vocabulary** words should also be highlighted in bold.

#### `code type-face`

- `Variable` names, `functions` and `arguments` of functions should be written in the code type-face.

```
code chunks
```

```
Any line of code that a student will run should be written as a code chunk.
```

#### _Italics_

- Italics should be used to add _emphasis_ to words or concepts.
- The _Panes_ and _Tabs_ within RStudio should also be italicized. 


### More on code chunks

#### Each individual line of code should be written as it's own code chunk. 

- Good
```{r, eval=FALSE}
x <- sample(1:10, size = 100, replace = )
```
```{r,eval=FALSE}
y <- 3*x - 2
```

- Bad
```{r, eval=FALSE}
x <- sample(1:10, size = 100, replace = )
y <- 3*x - 2
```

#### Function arguments should be written explicitly.
- Good
```{r, eval=FALSE}
seq(from = 0, to = 100, by = 10)
```

- Bad
```{r, eval=FALSE}
seq(0, 100, 10)
```    

#### Code chunks should be written to be as readable as possible on the slides.

- Good
```{r,eval=FALSE}
histogram(~ height | gender, 
          data = cdc, 
          col = "steelblue", 
          layout = c(1,2))
```

- Bad
```{r,eval=FALSE}
histogram(~height | gender, data = cdc, col = "steelblue", layout = c(1,2))
```

### Objects formatting

#### Objects should be named as concisely and descriptively as possible.

- Good
```{r, eval=FALSE}
cdc_males <- subset(cdc, gender == "Male")
```

- Bad
```{r, eval=FALSE}
df <- subset(cdc, gender = "Male")
```

- Bad
```{r, eval=FALSE}
all_male_subset_of_the_cdc_data <- subset(cdc, gender = "Male")
```

#### Objects should also be created using lowercase letters and underscores if necessary.

- Good
```{r, eval=FALSE}
cdc_females <- subset(cdc, gender == "Female")
```

- Bad 
```{r, eval=FALSE}
cdcFemales <- subset(cdc, gender == "Female")
```

- Bad 
```{r, eval=FALSE}
CDCFemales <- subset(cdc, gender == "Female")
```

- Bad 
```{r, eval=FALSE}
cdc.females <- subset(cdc, gender == "Female")
```

- Bad
```{r, eval=FALSE}
CDC_Females <- subset(cdc, gender == "Female")
```
