---
title: "What's the Relationship?"
author: "Lab 1E"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../mathjax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Finding patterns in data.

- To discover (_really_) interesting observations or relationships in data, we need to find them!
    - Which is difficult if we only look at the raw data.
- The best tool for finding patterns is often ... your own eyes.
    - Plots are an excellent way to help your eye search for patterns.
- In this lab, we'll learn how to include more variables in our plots to make them more informative.
- To begin, load the `cdc` data.


## Where's the variables?

```{r,eval=TRUE, echo=FALSE, fig.align='center', fig.width=14, fig.height= 6, out.width = '90%'}
histogram(~height | gender, data = cdc, layout = c(2,1),
          scales=list(cex=2.5),
          xlab = list("Height", cex=2.5),
          ylab = list("Count", cex=2.5),
          par.strip.text=list(cex=2))
```

- **How many variables were used to create this plot? Which variables were used and how were they used?**


## Multiple variable plots

- The previous graph is an example of a _multiple variable plot_, which means that more than a single variable was used. In this case:
  - Variable 1: _height_
  - Variable 2: _gender_
- Multiple variable plots are tools for finding _relationships_ between data.
- Let's take our `cdc` data and make some new multiple variable plots you haven't created before!


## scatterplots

```{r,echo=FALSE, fig.align='center', fig.width=8, fig.height=6, out.width = '60%'}
n = 50
x <- runif(n, 0, 20)
y <- x^3 + rnorm(n, mean = 0, sd = 10^3)
df <- data.frame(x=x, y=y)
xyplot(y~x, data = df,
       cex = 2,
       scales=list(cex=2.5),
          xlab = list("Variable 1", cex=2.5),
          ylab = list("Variable 2", cex=2.5))
```


## Creating scatterplots

- Scatterplots are useful for viewing how one _numerical_ variable relates to another _numerical_ variable.
- Fill in the blanks to create a scatterplot with `weight` on the y-axis and `height` on the x-axis.
```{r, eval = FALSE, echo = TRUE}
xyplot(____ ~ ____, data = cdc)
```


## Scatterplots in action

- Use a scatterplot to answer the following questions:
    - **As we look at people who are taller, what happens to their weights?**
    - **What happens if you swap the `weigth` and `height` variables in your code? Does the relationship between the variables change?**
    - **Does the relationship between `height` and `weight` change when we facet by gender? Write down the code you used to answer this question.**


## 4-variable scatterplots

- When we make scatterplots, we can include:
    - 1 numerical variable on the x-axis
    - 1 numerical variable on the y-axis
    - Use 1 categorical variable to facet our scatterplot
    - Change the color of the points based on another categorical variable
- To change the color of our points, we can include the `groups` argument  much like we did for bargraphs (use the _search_ feature in the _History_ pane if you need help).
- **Create a scatterplot that uses these 4 variables: `height`, `weight`, `gender`, `depressed`.**


## Multiple facets

- It can sometimes be helpful to facet on more than 1 variable.
    - Splitting the the data using 2 facets can give us additional insights that might otherwise be hidden.
- Create a `dotPlot` or `histogram` of the `weight` variable, but facet the data using:
```{r, eval = FALSE, echo = TRUE}
gender + depressed
```
- **How does the `gender` and feeling depressed impact a person's weight?**

