---
title: "Scraping web data"
author: "Lab 3E"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## The web as a data source

- The internet contains huge amounts of information.
  - Using computers to gather this information in an automated fashion is referred to as _scraping web data_.
  - Scraping data from the web can be difficult because each website displays & stores data differently.
- In this lab, we'll learn how to scrape data in two steps:
  - Step 1: Gather information from the web.
  - Step 2: Clean it up and turn it into a usable data frame for `Lab 3F`.
  

## Our first web scraper

- Copy and paste the link below into a web browser to view the website of data we'd like to _scrape_ and analyze.

[http://gh.idsucla.org/ids_labs/extras/webdata/mountains.html](http://gh.idsucla.org/ids_labs/extras/webdata/mountains.html)

- **Briefly describe what the data on the website is about.**
    - **Then write down 3 questions you'd be interested in answering by analyzing this data.**


## HTML

- `HTML` is the code that's used to render every website you've ever visited.
- The following slide shows the `HTML` code used to create the first two rows of the web data.
    - **How is the data table in `HTML` different than the data tables we're used to seeing in `R`, for example, when we use the View() function?**
    - **What do you think the _tags_ `<TABLE>`, `<TR>`, `<TH>`, `<TD>` mean? How does `HTML` use these _tags_ to display the table?**
    

## 

<font size = "6">
```
<TABLE>
  <TR>
    <TH>peak</TH>
    <TH>range</TH>
    <TH>state</TH>
    <TH>long</TH>
    <TH>lat</TH>
    <TH>elev_ft</TH>
    <TH>elev_m</TH>
    <TH>prominence_ft</TH>
    <TH>prominence_m</TH>
    <TH>rank</TH>
  </TR>
  <TR>
    <TD>Denali (Mount McKinley)</TD>
    <TD>Alaska Range</TD>
    <TD>Alaska</TD>
    <TD>-151.0063</TD>
    <TD>63.0690</TD>
    <TD>20236</TD>
    <TD>6168</TD>
    <TD>20174</TD>
    <TD>6149</TD>
    <TD>1</TD>
  </TR>
</TABLE>
```
</font>


## Get to scraping!

- Use your browser to go back to the website with the data we're interested in scraping.
- Find the URL address for the site and assign it the name `data_url` in `R`.
    - Then fill in the blanks below to have `R` scrape _every_ web table available on the site:

  ```{r, eval = FALSE, echo = TRUE}
  tables <- readHTMLTable(____)
  ```


## Find our data

- Since `readHTMLTable()` scrapes _every_ table that is on a particular web URL, we need to find out which table has the data we're interested in.
    - For example, `wikipedia.org` often has articles with 3 or more tables.
    - This means we need to check all 3 tables to find the one we're interested in.
- Use the `length()` function to find out how many tables of data were scraped in our set of `tables`.


## Saving tables

- Now that we know how many tables we've scraped, we can go back and scrape individual tables by adding the `which` argument to the `readHTMLTable()` function.
    - Use `readHTMLTable()` to re-scrape the data from the web but this time use the `which` argument to scrape just the individual table.
    - The `which` argument should be the integer denoting which table you want scraped.
    - Assign the scraped data the name `mtns`


## From scraping to cleaning

- Data scraped from the web usually needs to be cleaned.
- **Run the following commands and compare the names of the variables. Do you notice any differences?**
  ```{r, eval = FALSE, echo = TRUE}
  View(mtns)
  names(mtns)
  ```
- **Which variables in your data are numerical variables and which are factors (i.e. categorical variables)?**
- Put your data in the `str()` function to see how `R` classified each variable.
    - **Which variables are wrong?**  


## Fixing variable types

- View the `mtns` data and notice the order of the variables. 
    - Use the order of the variables to fill in the blanks below with either the word `"factor"`, if the variable is _categorical_, or `"numeric"`, if the varible is _numerical_.
  ```{r, eval = FALSE, echo = TRUE}
  var_types <- c("___","___","___","___","___",
                 "___","___","___","___","___")
  ```

- Finally, re-scrape the data and include `colClasses = var_types` as an argument.
    - Don't forget to save the data as `mtns` and specify `which` table to scrape.


## Fixing variable names

- View the `mtns` data and notice the order of the variables.
    - Then use the order of the variables and the following code template to change the names of the `mtns` data.
    - Replace each `"new_name"` with the actual name of the variable. 
    - Make sure to include all of the variable names and order them correctly.
  ```{r, eval = FALSE, echo = TRUE}
  names(mtns) <- c("new_name", "new_name",
                   ..., "new_name")
  ```


## Check, save and use!

- After scraping and cleaning the data, the only thing left to do is to save it and use it.
    - Before saving, use the `names()` and `str()` functions on last time to make sure the variable names and types are correct.
- Fill in the blanks to save the data and give it a file name
```{r, eval = FALSE, echo = TRUE}
save(____, file = "____.Rda")
```

- **What is the mean and standard deviation of `elev_ft`?**
- **Which `state` has the most mountains in our data?**
