---
title: "This model is big enough for all of us!"
author: "Lab 4F"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Building better models

- So far, in the labs, we've learned how to make predictions using the _line of best fit_, also known as _linear models_ or _regression models_.
- We've also learned how to measure our model's prediction accuracy by cross-validation.
- In this lab, we'll investigate the following question:

_Will including more variables in our model improve its predictions?_


## Divide & Conquer

- <span class="blue">Start by loading the `movie` data and split it into two sets (see Lab 4C for help).</span>
    - <span class="blue">A set named `training` that includes 75% of the data.</span>
    - <span class="blue">A set named `test` that includes the remaining 25%.</span>
    - <span class="blue">Remember to use `set.seed`.</span>
- <span class="blue">Create a linear model, using the `training` data, that predicts `gross` using `runtime`.</span>
    - <span class="blue">Compute the MSE of the model by making predictions for the `test` data.</span>
- **Do you think that a movie's `runtime` is the only factor that goes into how much a movie will make? What else might affect a movie's `gross`?**


## Including more info

- Data scientists often find that including more relevant information in their models leads to better predictions.
    - <span class="blue">Fill in the blanks below to predict `gross` using `runtime` and `reviews_num`.</span>
```{r, eval = FALSE, echo = TRUE}
lm(____ ~ ____ + ____, data = training)
```

- **Does this new model make more or less accurate predictions? Describe the process you used to arrive at your conclusion.**
- **Write down the code you would use to include a 3rd variable, of your choosing, in your `lm()`.**


## On your own

- **Write down which other variables in the `movie` data you think would help you make better predictions.** 
    - **Are there any variables that you think would not improve our predictions?**
- <span class="blue">Create a model for all of the variables you think are relevant.</span>
    - **Assess whether your model makes more accurate predictions for the `test` data than the model that included only `runtime` and `reviews_num`.**
- **With your neighbors, determine which combination of variables leads to the best predictions for the `test` data.**
