---
title: "What's the Relationship?"
author: "Lab 1E"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  word_document:
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Finding patterns in data.

- To discover (_really_) interesting observations or relationships in data, we need to find them!
    - Which is difficult if we only look at the raw data.
- The best tool for finding patterns is often ... your own eyes.
    - Plots are an excellent way to help your eye search for patterns.
- In this lab, we'll learn how to include more variables in our plots to make them more informative.
- Import the data from your class' _Food Habits_ campaign and name it `food`.


## Where's the variables?

```{r,eval=TRUE, echo=FALSE, fig.align='center', fig.width=14, fig.height=6, out.width = '90%'}
histogram(~height | gender, data = cdc, layout = c(2,1),
          scales=list(cex=2.5),
          xlab = list("Height", cex=2.5),
          ylab = list("Count", cex=2.5),
          par.strip.text=list(cex=2))
```

- **How many variables were used to create this plot? Which variables were used and how were they used?**


## Multiple variable plots

- The previous graph is an example of a _multiple variable plot_, which means that more than a single variable was used. In this case:
  - Variable 1: _height_
  - Variable 2: _gender_
- Multiple variable plots are tools for finding _relationships_ between data.
- Let's take our `food` data and make some new multiple variable plots you haven't created before!


## scatterplots

```{r,echo=FALSE, fig.align='center', fig.width=8, fig.height=6, out.width = '60%'}
n = 50
x <- runif(n, 0, 20)
y <- x^3 + rnorm(n, mean = 0, sd = 10^3)
df <- data.frame(x=x, y=y)
xyplot(y~x, data = df,
       cex = 2,
       scales=list(cex=2.5),
          xlab = list("Variable 1", cex=2.5),
          ylab = list("Variable 2", cex=2.5))
```


## Creating scatterplots

- Scatterplots are useful for viewing how one _numerical_ variable relates to another _numerical_ variable.
- Fill in the blanks to create a scatterplot with `sodium` on the y-axis and `sugar` on the x-axis.
```{r, eval = FALSE, echo = TRUE}
xyplot(____ ~ ____, data = food)
```


## Scatterplots in action

- Use a scatterplot to answer the following questions:
    - **Do snacks that have more `protein` also have more `calories`? Why do you think that?**
    - **What happens if you swap the `protein` and `calories` variables in your code? Does the relationship between the variables change?**
    - **Does the relationship between `protein` and `calories` change when the snack is either `Salty` or `Sweet`? Write down the code you used to answer this question.**


## 4-variable scatterplots

- When we make scatterplots, we can include:
    - 1 numerical variable on the x-axis
    - 1 numerical variable on the y-axis
    - Use 1 categorical variable to facet our scatterplot
    - Change the color of the points based on another categorical variable
- To change the color of our points, we can include the `groups` argument  much like we did for bargraphs (use the _search_ feature in the _History_ pane if you need help).
- **Create a scatterplot that uses these 4 variables: `sodium`, `sugar`, `cost`, `salty_sweet`.**


## Multiple facets

- It can sometimes be helpful to facet on more than 1 variable.
  - Splitting the data using 2 facets can give us additional insights that might otherwise be hidden.

Create a `dotPlot` or `histogram` of the `calories` variable, but facet the data using:
```{r, eval = FALSE, echo = TRUE}
healthy_level + salty_sweet 
```
**How does the `healthy_level` of a `Salty` or `Sweet` snack impact the number of calories in the snack?**

- Although we are treating `healthy_level` as a categorical variable, `R` recongizes it as a numerical variable. 
  - Use the `str` command to confirm
  - Notice that the faceted `histograms` or `dotPlots` do not have labels but rather tick-marks
  - You will have the opportunity to convert the `healthy_level` variable into a factor later on

- Faceting your data on a numerical variable is NOT recommended
  - Numerical variables often have so many different values that they overwhelm the plot and make it hard to read

## On your own

- Answer the following questions by creating an appropriate graph or graphs.
    - **Do healthier snacks have more or less `ingredients` than less healthy snacks?**
    - **What other variables seem to be related to the number of `ingredients` of a snack? Describe their relationships.**