---
title: "The Titanic Shuffle"
author: "Lab 2F"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE}
library(mobilizr)
data(titanic)
```

## Previously ...

- In the previous lab, we learned that by using a `do`-loop and the `shuffle` function, we could simulate randomly shuffling our data many times.
  - This helps us determine how likely it is that a difference between groups is due to chance.
- For this lab, we will extend these ideas to _numerical_ variables by using random shuffling and numerical summaries.
- The question we will investigate in this lab is:

_Is there any evidence to suggest that those who survived paid a higher fare than those who died?_


## The Titanic

- The Titanic was a ship that sank en route to the U.S.A. from England after hitting an iceberg in 1912.
    - At the time, it was claimed that the Titanic was _unsinkable_ ... it wasn't ... because it did.
- <span class="blue">Use the `data` function to load the `titanic` passenger and survival data.</span>
- <span class="blue">Create a boxplot of the `fare`s paid by passengers and facet the plot based on whether the passenger survived or not.</span>
    - **Based on the plot, do you believe that passengers who paid a higher fare on the Titanic were more likely to survive? Explain why and describe how certain you are of being correct.**


## The search begins!

- <span class="blue">Start your analysis by calculating how much more the _typical_ survivor paid than the _typical_ non-survivor in our data.</span>
- **Based on the distributions of fares paid, which numerical summary that describes the _typical_ value might be preferred?**
- **What was the _typical_ fare paid by survivors? Non-survivors? How much more did the typical survivor pay?**


## Do the shuffle!

- <span class="blue">Use the `do` and the `shuffle` functions to `shuffle` the passenger's survival status 500 times.</span>
    - Use the previous lab if you need some help on how to do this.
    - For each shuffle, compute each group's `median` fare paid. 
    - Assign your shuffled data the name `shuffled_survival`.
- <span class="blue">After shuffling your data, use the `mutate` function to create a variable called `diff` which is the `median` fare of survivors minus the `median` fare of non-survivors.</span>
    - Assign your mutated data the name `shuffled_survival` again.
  

## Put your simulations to use

- **Using your shuffled data, answer the research question we posed at the beginning of the lab.**

_Is there any evidence to suggest that those who survived paid a higher fare than those who died?_

- **Write up your answer as a statistical analysis. Create a plot and explain how the plot supports your conclusion. Be sure to also explain why shuffling your data is important.**


## Comparing Mean Fares

- What about if instead of calculating the median fare price for each group after a shuffle, we calculated the mean fare price and took the difference (mean_survivor – mean_victim)?

- **If we did this 500 times, what do you predict the distribution of differences will look like?**

- <span class="blue">Use the `do` and the `shuffle` functions to shuffle the passenger survival status 500 times. </span>
    - For each shuffle, compute each group's mean fare paid.
    - After shuffling your data, use the `mutate` function to create a variable called `diff` which is the mean fare of survivors minus the mean fare of non-survivors.
    
- **What does the shuffled data reveal? Does the answer to the research question below change when using the mean fares instead of the median fares?**

_Is there any evidence to suggest that those who survived paid a higher fare than those who died?_