---
title: "Base R Basics"
author:
date: The following slides contain codes to create graphics that can be utilized in the IDS course.
output:
  slidy_presentation: default
  pdf_document: default
---

```{r setup, include=FALSE}
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
library(readr)
# colors <- read_csv("colors.csv")
colors <- extra_data("colors_ids")
colors <- colors_ids
data(arm_span)
```

## The basic plots

Use the command below to load the colors_ids dataset:

```{r, eval=FALSE}
extra_data("colors_ids")
```


```{r, fig.height=4, fig.width=5}
hist(colors_ids$blue)
```


```{r, fig.height=4, fig.width=5}
boxplot(colors_ids$blue)
```

```{r, fig.height=4, fig.width=5}
barplot(table(colors_ids$p_color))
```


```{r, fig.height=4, fig.width=5}
plot(colors_ids$blue, colors_ids$gold)
```

## Arguments


Many of the arguments that work for the lattice graphics transfer to base R.

- `main`
- `col`
- `ylab`
- `xlab`
- `pch` for scatterplots
- `border` for bars

## Split histogram

There are times when you facet a histogram and you would like each histogram to be a different color. The code below shows how to accomplish this:

```{r}
##personality color
# Separate colors into vectors
green <- colors_ids$green
blue <- colors_ids$blue
gold <- colors_ids$gold
orange <- colors_ids$orange

# Set up margins (for Error: margins are too large)
#par sets the parameters for graphs
#mar sets the margins
#the c() function is creating a vector for the bottom, left, top, right margins
par(mar = c(2, 1, 1, 1) + 0.1)

# Set up the layout
layout(matrix(1:4, nrow = 4, byrow = TRUE))

# Create stacked histograms for each personality color
hist(green, col = "seagreen3", 
     main = "Green", 
     xlab = "Score", 
     border = "white",
     breaks = seq(5, 20, by=1))

hist(blue, col = "royalblue1", 
     main = "Blue", 
     xlab = "Score", 
     border = "white",
     breaks = seq(5, 20, by=1))

hist(gold, col = "gold", 
     main = "Gold", 
     xlab = "Score", 
     border = "white",
     breaks = seq(5, 20, by=1))

hist(orange, col = "orange", 
     main = "Orange", 
     xlab = "Score", 
     border = "white",
     breaks = seq(5, 20, by=1))

# Reset the layout for your next plot
layout(1)
```


## Some useful functions

```{r, fig.height=4, fig.width=5}
plot(arm_span$armspan, arm_span$height)
#Points function in base R is more user friendly
points(60, 62.6667, col = "red", cex = 2)
points(64, 64, col = "red", cex = 2)
points(68, 68, col = "red", cex = 2)
points(72, 71.5, col = "red", cex = 2)
```



## On Your Own

If you are looking for some new datasets, check out our available pre-loaded data sets by running the command:

```{r, eval=FALSE}
data()
```

Select a pre-loaded data set of your choice, and create the four basic plots using the dollar sign syntax. Include arguments to make your graphic clear.






