---
title: "Base R Basics"
author:
date: The following slides contain codes to create graphics that can be utilized in the IDS course.
output:
  slidy_presentation: default
  pdf_document: default
---

```{r setup, include=FALSE}
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
library(readr)
# colors <- read_csv("colors.csv")
colors <- extra_data("colors_ids")
colors <- colors_ids
data(arm_span)
```

## The basic plots

```{r, fig.height=4, fig.width=5}
hist(colors_ids$blue)
```


```{r, fig.height=4, fig.width=5}
boxplot(colors_ids$blue)
```

```{r, fig.height=4, fig.width=5}
barplot(table(colors_ids$p_color))
```


```{r, fig.height=4, fig.width=5}
plot(colors_ids$blue, colors$gold)
```

## Arguments


Many of the arguments that work for the lattice graphics transfer to base R.

- `main`
- `col`
- `ylab`
- `xlab`
- `pch` for scatterplots

## Some useful functions

```{r, fig.height=4, fig.width=5}
plot(arm_span$armspan, arm_span$height)
#Points function in base R is more user friendly
points(60, 62.6667, col = "red", cex = 2)
points(64, 64, col = "red", cex = 2)
points(68, 68, col = "red", cex = 2)
points(72, 71.5, col = "red", cex = 2)
```



## On Your Own

If you are looking for some new datasets, check out our available pre-loaded data sets by running the command:

```{r, eval=FALSE}
data()
```

Select a pre-loaded data set of your choice, and create the four basic plots using the dollar sign syntax. Include arguments to make your graphic clear.






