---
title: "ggplot 2 basics"
author:
date: The following slides contain codes to create cool graphics that can be utilized in the IDS course.
output:
  slidy_presentation: default
  pdf_document: default
---

```{r setup, include=FALSE}
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
library(readr)
library(ggplot2)
# colors <- read_csv("colors.csv")
colors <- extra_data("colors_ids")
colors <- colors_ids
```

## Boxplots and color scores

Use the command below to load the colors_ids dataset:

```{r, eval=FALSE}
extra_data("colors_ids")
```


```{r}
ggplot(data=colors_ids)+
  geom_boxplot(mapping=aes(x=blue), fill="blue")+
  #need to change the value of y so plots don't overlap
  geom_boxplot(mapping=aes(x=orange, y=1), fill="orange")+
  geom_boxplot(mapping=aes(x=green, y=2), fill="green")+
  geom_boxplot(mapping=aes(x=gold, y=3), fill="gold")+
  #show values on the x-axis
  scale_x_continuous(breaks=seq(5,20,1))+
  labs(title="Distribution of color scores",x="scores", y="")+
  theme_bw()+
  #hides the y-axis labels
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

## Boxplots and dotplots

```{r}
ggplot(data=colors_ids)+
  geom_boxplot(mapping=aes(x=blue, y=1))+
  scale_x_continuous(breaks=seq(5,20,1))+
  labs(title="Distribution of blue scores")+
  theme_bw()+
  geom_dotplot(mapping=aes(x=blue),fill="blue",binwidth = 0.5)
```

## Bargraphs and faceting

```{r}
ggplot(data=colors_ids)+
  geom_bar(mapping=aes(x=p_color, fill=p_color))+
  scale_fill_manual(values=c("blue", "gold", "green","orange"))+
  theme_classic()
```


```{r}
ggplot(data=colors_ids)+
  geom_bar(mapping=aes(x=p_color, fill=p_color))+
  scale_fill_manual(values=c("blue", "gold", "green","orange"))+
  theme_classic()+
  facet_grid(rows=vars(int_ext))
  #replace `rows` with `cols` above
```
## Scatterplots

The scatterplot below has 90 data points but because some of the armspans and corresponding heights overlap there appear to be less.


```{r}
ggplot(data=arm_span)+
  geom_point(mapping=aes(x=armspan, y= height))
```

A solution to this problem is to utilize the `geom_jitter` function. 


```{r}
ggplot(data=arm_span)+
  geom_point(mapping=aes(x=armspan, y= height))+
  geom_jitter(mapping=aes(x=armspan, y= height))
```

Alternative code

```{r, eval=FALSE}
ggplot(data=arm_span, mapping=aes(x=armspan, y= height))+
  geom_point()+
  geom_jitter()
```


## On Your Own

Choose one of the plots from this lab and make at least 3 modifications to it.

Here are some other themes you can choose from:

  - theme_gray()
  - theme_dark()
  - theme_light()
  - theme_linedraw()
  - theme_minimal()
  - theme_void()
  
