---
title: "Enhancing Lattice Graphics"
author:
date: The following slides contain **new** commands that can be used to enhance the lattice graphics used in the IDS Course.
output:
  slidy_presentation: default
  pdf_document: default
---

```{r setup, include=FALSE}
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
library(readr)
library(RColorBrewer)
```

## Let's get our data

-   We usually load datasets by using the command `data()`, but for this lab, we are going to read in your data using the commands below:

Copy and paste these commands into your console and press enter:

```{r, eval=FALSE}
utah<-read.csv("https://raw.githubusercontent.com/stedy/Machine-Learning-with-R-datasets/master/insurance.csv")
```

You should see a dataframe called "utah" in your `Environment Pane`.

## Categorical Variables

A plot that is appropriate to analyze categorical data is a `bargraph`.

The command that `R` recognizes to make a bargraph IS `bargraph`. There are two pieces of information, or arguments, that you need to provide to `R` to make a bargraph:

- variable
- dataframe

Syntax is extremely important so let's look and see how the variables were named by running the command below:

```{r, eval = F}
names(utah)
```

Now we are ready to make our `bargraph`.
Replace the word `variable` with one of the names of a categorical variable and run the code.

```{r, eval=F}
bargraph(~variable, data = utah)
```

- What is this bargraph telling you?

## More Arguments

Although the `bargraph` function only needs two arguments to make a bargraph, there are additional arguments that can be used to increase the readability of your graph. 

Let's add a title by adding the argument main.

```{r, eval = F}
bargraph(~variable, data = utah, main = "Here is your title")
```


## Numerical Variables

There are many plots that can be used to analyze the distribution of a single numerical variable such as a histogram, dotplot, and boxplot. Since we have a small dataset to work with, we will use a dotplot.

The command that `R` recognizes to make a dotplot is `dotPlot`(note the capital P). You need the same two arguments to make a `dotPlot`.

Use what you have learned so far, to create a `dotPlot` of a numerical variable. (You are engaging in computational thinking!)





