---
title: "Getting It Together"
author: "Lab 2G"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output:
  slidy_presentation:
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
  word_document:
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE}
library(mobilizr)
```

## Putting data together

- In the labs so far, we've only ever looked at individual data files.
  - But often times, we gain additional insights by including additional information from a separate dataset.
- In this lab, we will learn how to merge information from our _personality color_ data with our _stress/chill_ data.
- <span class="blue">_Export, upload, import_ your _Personality Color_ dataset and name it `colors`.</span>
- <span class="blue">Then, _export, upload, import_ your _Stress/Chill_ dataset and name it `stress`.</span>


## Looking at Stress/Chill

- We would like to analyze the research question:

_How do people's personality colors and/or sports participation affect their stress levels?_

- We already have data about _personality color_ and a separate dataset about _stress_.
    - What we don't have is a single dataset with information from both ... yet.
- We'll start then by strategizing how to merge our data together.

    
## Deciding how to merge

- Before we merge data, we need to decide _how_ we plan to merge it:
- We can _stack_ our datasets, that is, take one dataset's rows and add them to the bottom of the other dataset.
- We can also _join_ our datasets horizontally. This is where we take one dataset's columns and add them to the end of the other dataset's columns based on matching an _ID_ variable.
    - The _ID_ variable will have entries that we use to _match_ observations in both datasets.
- **To answer the statistical question of interest, would it make more sense to _stack_ or _join_ our `colors` and `stress` data?**


## Finding variables in common:

- Look at the `names` of the variables in each dataset.
    - To merge different datasets together, we need to find variables they have in common.
- **Which variables do the datasets have in common?**
- **Which variable would make sense to merge the datasets together with? Why not the others?**
 

## Caution required

- Whether _stacking_ or _joining_, we need to be careful when we merge data:
- When _stacking_ data, we need to be absolutely certain that the variables we're stacking represent the exact same measurements.
    - We wouldn't want to stack `height` in meters and `height` in inches, for instance (without converting one to the other).
- When _joining_ data, we need to make sure that the _id_ variable in our primary dataset matches to _one and only one_ observation in the joining data.
    - Otherwise, `R` won't know which observation to match to.
    

## Getting ready

- Our goal is to add the variables from the `colors` data onto the `stress` data.
- <span class="blue">Start by ensuring that every `user.id` in the `colors` data is unique.
    - If there's a duplicate, have your teacher remove the duplicate from your class' Web Response Manager and then re-_export_, _upload_, _import_ your `colors` data.</span>
- **After we add the data from _colors_ to _stress_, how many rows should our merged data have? Write this number down.**


## Putting them together

- We can use the `merge` function to _join_ our datasets together using the variables that appear in both sets.
- <span class="blue">Fill in the blanks below to join the information from the `colors` data onto the `stress` data.</span>
```{r, eval = FALSE, echo = TRUE}
merge(____, ____, by = "____")
```
- <span class="blue">`Assign` this `merged` dataset the name `stress_colors`.</span>
    - Make sure your data has the same number of observations that you wrote down on the previous slide.


## Saving your data:

- <span class="blue">`View` your merged data and make sure nothing appears to be blatantly wrong with it.</span>
- **Why didn't we stack the rows of data instead?**
- **What happens if you swap the order of the datasets in the `merge` function?**
- <span class="blue">Fill in the blank below to `save` our `stress_colors` data for later use.</span>
    
```{r, eval = FALSE, echo = TRUE}
save(stress_colorss, file = "stress_colorss.Rda")
```

- <span class="blue">Be sure to look in the _Files_ tab to make sure your data was saved.</span>


## Moving on

- In the next practicum, we'll begin analyzing our merged data. In the meantime: 
- <span class="blue">Make a few plots using variables from the `stress` data and _facet_ or _group_ the plots based on variables from the `colors` data.</span>
- **Write down the most interesting discovery you make by just exploring your data. Write out how you found your discovery and interpret what it means for the people in your class.**
- **With our _colors_ data, we could answer questions about the _typical_ color scores in your class. Why can we no longer answer this question in our `stress_colors` data?**
