---
title: "Queue it up!"
author: "Lab 2D"
date: "Directions: Follow along with the slides and answer the questions in **red** font in your journal."
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
 
```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Where we left off

- In the last lab, we looked at how we can use computer simulations to compute estimates of simple probabilities.
  - Like the probability of drawing a song genre from a playlist.
- We also saw that performing _more_ simulations:
    - Took _longer_ to finish.
    - Had estimates that _varied less_.
- In this lab, we'll extend our simulation methods to cover situations that are more complex.
    - We'll learn how to estimate their probabilities.
    - We also look at the roll of sampling _with_ or _without_ _replacement_.
    

## Back to songs

- In `R`, simulate a _playlist of songs_ containing 30 `"rap"` songs, 23 `"country"` songs and 47 `"rock"` songs.
    - _Assign_ the `c`ombined playlist the name `songs`.
- Simulate choosing a single song 50 times. Then use your simulated draws to estimate the probability of choosing a _rap_ song.
    - The actual (theoretical) probability of choosing a _rap song_ in this case is `0.30`.
    - **Write a sentence comparing your estimated probability to the actual probability.**


## With or Without?

- So far, you've selected songs _with replacement_. 
    - We called it that, because each time you made a selection, you started with the same playlist. That is, you chose a song, wrote down its data, and then placed  it back on the list.
- It's also possible to select _without replacement_ by setting the `replace` option in the `sample` function to `FALSE`.
- Take a sample of `size` 100 from our playlist of songs _without replacement_. Assign this sample the name `without`. 
    - **What do you notice if you run `tally(~without)`? Does something similar happen if you sample _with replacement_?**
    - **What happens if `size = 101` and `replace = FALSE`?**


## Sample with? Or without?

- Imagine the following two scenarios.
    1. You have a coin with two sides: _Heads_ and _Tails_. You're not sure if the coin is fair and so you want to estimate the probability of getting a _Head_.
    2. A child reaches into a candy jar with 10 _strawberry_, 50 _chocolate_ and 25 _watermelon_ candies. The child is able to grab three candies with their hand and you're interested in probability that all three candies will be chocolate.
- **Which of these scenarios would you sample _with replacement_ and which would you sample _without replacement_? Why?**
    - **Write down the line of code you would run to `sample` from the candy jar. Assume the simulated jar is named `candies`.**


## Simulations at work

- In reality, songs from a playlist are chosen without replacement.
    - This way, you won't hear the same song several times in a row.
- Let's write a more realistic simulation and estimate the probability that if we select two songs at random, without replacement, that both are rap songs.
    - Use the `do` function to perform 10 simulated `sample`s of `size` 2, with replacement and _assign_ the simulations the name `draws`.
    

## Simulations and probability

- To estimate the probability from our simulations, we need to find the proportion of times that the event we're interested in occurs in the simulations.
- In other words, we need to count the number of times the desired events occurred, divided by the number of attempts we made (the number of simulations).
- The next slides will show you two ways to do this.


## Counting similar outcomes

- One way we can estimate the probability of drawing two songs of the _same_ genre is to use the following trick to count the number of _rap_ songs in each of the 10 simulations:

```{r, eval = FALSE, echo = TRUE}
mutate(draws, nrap = rowSums(draws=="rap"))
```

- **For each of the lines of code below, describe how the output of the code changes as we move from line to line.**
```{r, eval = FALSE, echo = TRUE}
draws == "rap"
```
```{r, eval = FALSE, echo = TRUE}
rowSums(draws == "rap")
```
```{r, eval = FALSE, echo = TRUE}
mutate(draws, nrap = rowSums(draws=="rap"))
```


## Counting other outcomes

- Another method we can use to estimate the probability of complex events is to use the following 2-step procedure:
    1. Subset the rows of the simulations that match our desired outcomes.
    2. Count the number of rows in the subset and divide by the number of simulations.
- The result that you obtain is an estimate of the probability that a specific combination of events occured.
- We'll see an example of this method on the next slide.


## Step 1: Creating a subset

- Fill in the blanks below to: 
    1. Create a subset of our simulations when both draws were `"rap"` songs.
    2. Count the number of rows in this subset
    3. And divide by the total number of repeated simulations.
    
```{r, eval = FALSE, echo = TRUE}
draws_sub <- filter(draws, ___ == "rap",  ___ == "rap")
```
```{r, eval = FALSE, echo = TRUE}
nrow(___) / ___
```


## Estimating probabilities

- **Calculate estimated probabilities for the following situations:**
    1. You draw two `"rap"` songs.
    2. You draw a `"rap"` song in the first draw and a `"country"` song in the 2nd.
- **Create a histogram that displays the number of times a `"rap"` song occurred in each simulation. That is, how often were zero rap songs drawn? A single rap song? Two rap songs?**


## On your own

- Using what you've learned in the previous two labs, answer the following question by performing two computer simulations with 500 repetitions a piece:

**_If we draw 5 songs from a playlist of 30 rap, 23 country and 47 rock songs, how does the estimated probability of all 5 songs being rap songs change if we draw the songs with or without replacement?_**

- For each simulation:
    - **Create a histogram for the number of _rap_ songs that occurred for each of the 500 repetitions.**
- **Describe how the distribution of the number of _rap_ songs changes depending on if we use replacement or not.**

