---
title: "If the line fits ..."
author: "Lab 4A"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## How to make predictions

- Anyone can make predictions.
    - Data scientists use data to inform their predictions by using the information learned from the sample to make predictions for the whole population.
- In this lab, we'll learn how to make predictions by finding the _line of best fit_.
    - You will also learn how to use the information from one variable to make predictions about another variable.


## Predicting heights

- <span class="blue">Use the `data()` function to load the `arm_span` data.</span>
- This data comes from a sample of 90 people in the Los Angeles area.
    - The measurements of `height` and `armspan` are in inches.
    - A person's `armspan` is the maximum distance between their fingertips when they spread their arms out wide.
- <span class="blue">Make a plot of the `height` variable.</span>
    - **If you had to predict the height of someone in the Los Angeles area, what single height would you choose and why?**
    - **Would you describe this as a _good_ guess? What might you try to improve your predictions?**


## Predicting heights knowing arm spans

- <span class="blue">Create two subsets of our `arm_span` data: 
    - One for `armspan >= 61` and `armspan <= 63`. 
    - A second for `armspan >= 64` and `armspan <= 66`.</span>
- <span class="blue">Create a `histogram` for the `height` of people in each subset.</span>
- **Answer the following based on the data:**
    - **What `height` would you predict if you knew a person had an `armspan` around 62 inches?**
    - **What `height` would you predict if you knew a person had an `armspan` around 65 inches?**
    - **Does knowing someone's `armspan` help you predict their `height`? Why or why not?**




<!-- Problems with using the mean -->
<!-- === -->

<!-- - Using `mean` values to make predictions has a few drawbacks which we highlight below: -->
<!-- - Calculate the `mean` `height` of people who have an `armspan` of 69 and 70 inches.  -->
<!--     - **How much taller is the person with a 70 inch `armspan` than the 69 inch `armspan`, on average? Does this seem reasonable?** -->
<!--     - **How tall would you predict someone who had a 69.5 inch `armspan` to be?** -->


## Fitting lines

<!-- - Making predictions using means can lead to drastic jumps in predicted values. -->
<!--     - They also leave us unable to make predictions for inbetween values, like 69.5. -->
<!-- - For these reasons, data scientists often _fit_ lines to their data to make predictions. -->
- Notice that there is a trend that people with a larger `armspan` also tend to have a larger mean `height`.
    - One way of describing this sort of trend is with a line.
- Data scientists often _fit_ lines to their data to make predictions.    
    - What we mean by _fit_ is to come up with a line that's close to as many of the data points as possible.
- <span class="blue">Create a scatterplot for `height` and `armspan`. Then run the following code.</span>
```{r, eval = FALSE, echo = TRUE}
add_line()
```

- <span class="blue">On the _Plot_ pane, click two data points to draw a line through.</span>

## Predicting with lines

- **Draw a line that you think is a good _fit_ and write down its equation. Using this equation:**
    - **Predict how tall a person with a 62 inch `armspan` and a person with a 65 inch `armspan` would be.**
    
<!-- - **How much taller is the person with a 70 inch `armspan` than the 69 inch `armspan`? Does this difference seem more reasonable than the difference we calculated using the means?**   -->
- Using a line to make predictions also lets us make predictions for `armspan`s that aren't in our data.
    - **How tall would you predict a person with a 63.5 inch `armspan` to be?**
- **Compare your answers with a neighbor. Did both of you come up with the same equation for a line? If not, can you tell which line fits the data best?**


## Regression lines

- If you were to go around your class, each student would have created a different line that they feel _fit_ the data best.
    - Which is a problem because everyone's line will make slightly different predictions.
- To avoid this variation in predictions, data scientists will use _regression lines_.
    - This line connects the mean `height` of people with similar `arm_spans`.
    - <span class="blue">Fill in the blanks below to create a _regression line_ using `lm`, or _linear model_:</span>

```{r, eval = FALSE, echo = TRUE}
lm(____ ~ ____, data = arm_span)
```


## Predicting with regression lines

- **Use the output of the code from the previous slide to write down the equation of the _regression line_ in the form**
<center>`y = a + bx`.</center>
- <span class="blue">Add this line to a scatterplot by filling in the blanks below:</span>
```{r, eval = FALSE, echo = TRUE}
add_line(intercept = ____, slope = ____)
```

- **Predict the `height` of a person with a 63.5 inch `armspan` and compare it with a neighbor. Ensure you both arrive at the same predicted value.**
- **Measure your `armspan` and use the regression line to predict your `height`. How close was the prediction?**
