---
title: "Interpreting correlations"
author: "Lab 4D"
date: 'Directions: Follow along with the slides, completing the questions in <font color=#25679E><b>blue</b></font> on your computer, and answering the questions in **red** in your journal.  
  
  <br>
  <br>
  Space, Click, Right Arrow or swipe left to move to the next slide.'
output: 
  slidy_presentation: 
    css: ../../IDSLabCSSRev.css
    mathjax: ../../extras/mathjax-local/MathJax.js
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval=TRUE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(mobilizr)
```

## Some background...

- So far, we’ve learned about measuring the success of a model based on how close its predictions come to the actual observations. 
- The _correlation coefficient_ is a tool that gives us a fairly good idea of how these predictions will turn out without having to make predictions on future observations.
- For this lab, we will be using the `movie` data set to investigate the following question:

_Which variables are better predictors of a movie's `audience_rating` when the predictions are made using a line of best fit?_


## Correlation coefficients

- The _correlation coefficient_ describes the _strength_ and _direction_ of the linear trend.
- It's only useful when the trend is linear and both variables are numeric.
```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height = 5, fig.width = 12, out.width = '90%'}
xyplot(audience_rating~critics_rating, data = movie,
       cex = 2,
       scales=list(cex=2.5),
          xlab = list("Variable 1", cex=2.5),
          ylab = list("Variable 2", cex=2.5))
```

- **Are these variables linearly related? Why or why not?**


## Correlation review I

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height = 5, fig.width = 12, out.width = '90%'}
xyplot(audience_rating~critics_rating, data = movie,
       cex = 2,
       scales=list(cex=2.5),
          xlab = list("Variable 1", cex=2.5),
          ylab = list("Variable 2", cex=2.5))
```

- Correlation coefficients with values close to 1 are very strong with a positive slope. Values close to -1 means the correlation is very strong with a negative slope. 
- **Does this plot have a positive or negative correlation?**


## Correlation review II

```{r, eval=TRUE, echo=FALSE, fig.align='center', fig.height = 4.5, fig.width = 12, out.width = '90%'}
xyplot(audience_rating~critics_rating, data = movie,
       cex = 2,
       scales=list(cex=2.5),
          xlab = list("Variable 1", cex=2.5),
          ylab = list("Variable 2", cex=2.5))
```

- Recall that if there is no linear relation between two numerical variables, the correlation coefficient is close to 0. 
- **What do you guess the correlation coefficient will be for these two variables?**


## The movie data

- <span class="blue">Load the `movie` data using the `data` command.</span>
- The data comes from a variety of sources like _IMDB_ and _Rotten Tomatoes_. 
    - The `critics_rating` contains values between 0 and 100, 100 being the best.
    - The `audience_rating` contains values that range between 0 and 10, 10 being the best.
    - `n_critics` and `n_audience` describe the number of reviews used for the ratings.
    - `gross` and `budget` describe the amount of money the film made and took to make.


## Calculating correlation coefficients!

- We can use the `cor()` function to find the particular correlation coefficient of the variables from the previous plot, which happen to be `audience_rating` and `critics_rating`.
- But note, the `cor()` function removes any observations which contains an `NA` value in either variable.
- <span class="blue">Calculate the correlation coefficient for these variables using the `cor` function. The inputs to the functions work just like the inputs of the `xyplot` function.</span>


## Now answer the following.

- **What was the value of the correlation coefficient you calculated?**
- **How does this actual value compare with the one you estimated previously?**
- **Does this indicate a strong, weak, or moderate association? Why?**
- **How would the scatter plot need to change in order for the correlation to be stronger?**
- **How would it need to change in order for the correlation to be weaker?**


<!-- What if we changed the data? -->
<!-- === -->
<!-- - **Would the correlation coefficient change if we took all of our values and subtracted 50 from all of the `critics_score`? Why do you think this is?** -->
<!-- - Test your answer by running the following code: -->
<!-- ```{r, eval=FALSE} -->
<!-- cor(audience_rating~(critics_rating-50),  -->
<!--      data = movie) -->
<!-- ``` -->
<!-- - **What value was returned for the correlation coefficient?** -->
<!-- - **How does this actual value compare with the one you estimated previously?** -->

<!-- What if we changed the data, again? -->
<!-- === -->

<!-- - **What happens to the correlation coefficient if we multiplied all of the `critics_rating` values by 50?  How do you explain this?** -->
<!-- - Test your answer by running the following code: -->
<!-- ```{r, eval=FALSE} -->
<!-- cor(audience_rating~(critics_rating*50), data = movie) -->
<!-- ``` -->
<!-- - **What value was returned for the correlation coefficient?** -->
<!-- - **How does this actual value compare with the one you estimated previously?** -->


## Correlation and Predictions

- <span class="blue">Find the two variables that look to have the strongest correlation with `critics_rating`.</span>
    - <span class="blue">Compute the correlation coefficients for `critics_rating` and each of the two variables.</span>
    - **Use the correlation coefficient to determine which variable has a stronger linear relationship with `critics_rating`.**
- <span class="blue">Fit two `lm` models to predict `critics_rating` with each variable and compute the MSE for each.</span> 
    - **Use the MSE to determine which variable is a better predictor of `critics_rating`.**
- **How are the correlation coefficient and the MSE related?**


## On your own

- <span class="blue">Select two different numerical variables from the `movie` data. Plot the variables using the `xyplot()` function.</span>
    - **Would calculating a correlation coefficient for the two variables be appropriate? Justify your answer.**
    - **Predict what value you think the correlation coefficient will be. Compare this value to the actual value. Finally, interpret what the actual correlation coefficient means.**
- **Work with your classmates to determine which two variables have the strongest correlation coefficient.**
  - **Why do you think these variables are so strongly related? Is using the correlation coefficient to describe the relationship appropriate and why/why not?**
